---
title: "Compare Treatments KRN 2019"
author: "Emilio A. Laca"
date: "6/17/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

# THIS SCRIPT ANALYZES THE DIFFERENCES BETWEEN UNITS THAT WERE DISKED OR NOT AND FLOODED OR NOT.
# AN HTML REPORT IS SUPPOSED TO BE PRODUCED.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Effect of action in previous 3 years

library(tidyverse)
library(car)
library(emmeans)
library(rsample)
library(MASS)
library(ggforce)
library(knitr)
library(kableExtra)
library(effects)
library(brms)
library(rstanarm)
library(BayesPostEst)
library(xtable)
library(coda)
library(ggmcmc)
library(ggridges)
library(tidybayes)
library(magrittr)
library(modelr)
library(nlme)
library(posterior)

```

## Data input and exploration
```{r dataExplore}

qdt_mass_g_m2_2021 <- read.csv("qdt_mass_g_m2_2021.txt") %>%
  dplyr::select(-X) %>%
  dplyr::filter(LIT == "MDC")

act <- read.csv("WHAP2021_PP_20220922_MUs_MDC_Actions.csv",
                na.strings=c(""," ","NA"), stringsAsFactors=FALSE) %>%
  mutate(subunit_ID = paste(LIT, Unit_Name, Subunit_Name, sep = "_"),
         tilled = factor(Action_yn),
         Action = replace_na(Action, "none")) %>%
  dplyr::filter(LIT == "MDC") %>%
  dplyr::select(LIT,
                Unit_Name,
                Subunit_Name,
                subunit_ID,
                tilled,
                Action)

# effects of MSM

water <- read.csv("Modoc_Flooding_Schedule.csv") %>%
  mutate(LIT = "MDC",
         Subunit_Name = SubUnit) %>%
  dplyr::select(LIT,
         Subunit_Name,
         Schedule,
         starts_with("y"))

qdt_mass_act_fld <- left_join(qdt_mass_g_m2_2021, act) %>%
  left_join(water) %>%
  dplyr::select(LIT,
                Unit_Name,
                Subunit_Name,
                subunit_ID,
                tilled,
                Action,
                y.MSM,
                y20178,
                y20189,
                y201920,
                y202021,
                st_g_m2,
                Schedule,
                st_mass_m2_var,
                wg_g_m2,
                wg_mass_m2_var,
                sw_g_m2,
                sw_mass_m2_var,
                tot_g_m2,
                tot_mass_m2_var) %>%
  drop_na() %>%
  dplyr::filter(tot_mass_m2_var != 0, Schedule != "Irreg") %>%
  mutate(wt_tot = 1 / sqrt(tot_mass_m2_var),
         wt_tot = length(tot_g_m2) * wt_tot / sum(wt_tot))

# qdt_mass_act <- left_join(qdt_mass_g_m2_2021, act) %>%
#   dplyr::select(LIT,
#                 Unit_Name,
#                 Subunit_Name,
#                 subunit_ID,
#                 tilled,
#                 Action,
#                 st_g_m2,
#                 st_mass_m2_var,
#                 wg_g_m2,
#                 wg_mass_m2_var,
#                 sw_g_m2,
#                 sw_mass_m2_var,
#                 tot_g_m2,
#                 tot_mass_m2_var) %>%
#   drop_na() %>%
#   dplyr::filter(tot_mass_m2_var != 0) %>%
#   mutate(wt_st = 1 / sqrt(st_mass_m2_var),
#          wt_st = length(st_g_m2) * wt_st / sum(wt_st)) %>%
#   mutate(wt_wg = 1 / sqrt(wg_mass_m2_var),
#          wt_wg = length(wg_g_m2) * wt_wg / sum(wt_wg)) %>%
#   mutate(wt_sw = 1 / sqrt(sw_mass_m2_var),
#          wt_sw = length(sw_g_m2) * wt_sw / sum(wt_sw)) %>%
#   mutate(wt_tot = 1 / sqrt(tot_mass_m2_var),
#          wt_tot = length(tot_g_m2) * wt_tot / sum(wt_tot))

str(qdt_mass_act)

```

The raw data plots seems to indicate that the history of flooding is not associated with swamp timothy productivity. Conversely, those units that were disced in the previous three years had higher yields of swamp timothy than the units that were not tilled. 

```{r rawPlots}

table(qdt_mass_act_fld$Action) %>%
  as.data.frame() %>%
  rename(Tillage = Var1) %>%
  kable(caption = "Number of units that receive each type of tillage treatment in Kern prior to the 2021 season.") %>%
  kableExtra::kable_styling(full_width = FALSE)

scatterplot(tot_g_m2 ~ y.MSM | tilled,
            data = qdt_mass_act_fld,
            smooth = FALSE,
            ylab = "Productivity (g/m2)",
            xlab = "Number of years of MSM in the last 4 years")

```

## Analyze **apparent** effects of "treatments"


These analyses should be interpreted with caution. Apparent effects of treatments cannot be attributed to treatments because there was no randomization in the selection of units to be flooded. Based on my understanding of management decisions, flooding is decided on the basis of the potential of units to express higher production. Therefore, I am suspicious that there may be a selection bias in the results.


```{r posteriorTillageEffects}

brm1 <- brm(tot_g_m2 | weights(wt_tot) ~ tilled * y.MSM,
            data = qdt_mass_act_fld,
            iter = 4000,
            warmup = 1500,
            thin = 2,
)

summary(brm1)

bayes_R2(brm1)

brm1.mcmc <- as.mcmc(brm1)

# Check convergence, etc.

gelman.diag(brm1.mcmc)

```

### Graphical tests

```{r density}

# hh_m1_gg <- ggs(hh_m1_mcmc)

brm1.mcmc %>%
  ggs() %>%
  ggs_density(hpd = TRUE) +
  facet_wrap(~ Parameter, scales = "free")


```


```{r trace}

brm1.mcmc %>%
  ggs() %>%
  ggs_traceplot() +
  facet_wrap(~ Parameter, scales = "free")

```


```{r running}

brm1.mcmc %>%
  ggs() %>%
  ggs_running() +
  facet_wrap(~ Parameter, scales = "free")

```


```{r acf}

brm1.mcmc %>%
  ggs() %>%
  ggs_autocorrelation() +
  facet_wrap(~ Parameter, scales = "free")

```


### Reduced model

```{r posteriorTillageEffects2}

brm2 <- brm(tot_g_m2 | weights(wt_tot) ~ tilled,
            data = qdt_mass_act,
            iter = 6000,
            warmup = 2500,
            thin = 2)

summary(brm2)

bayes_R2(brm2)

brm2.mcmc <- as.mcmc(brm2)

# Check convergence, etc.

# gelman.diag(brm2.mcmc)

```

### Graphical tests

```{r density2}

# hh_m1_gg <- ggs(hh_m1_mcmc)

brm2.mcmc %>%
  ggs() %>%
  ggs_density(hpd = TRUE) +
  facet_wrap(~ Parameter, scales = "free")


```


```{r trace2}

brm2.mcmc %>%
  ggs() %>%
  ggs_traceplot() +
  facet_wrap(~ Parameter, scales = "free")

```


```{r running2}

brm2.mcmc %>%
  ggs() %>%
  ggs_running() +
  facet_wrap(~ Parameter, scales = "free")

```


```{r acf2}

brm2.mcmc %>%
  ggs() %>%
  ggs_autocorrelation() +
  facet_wrap(~ Parameter, scales = "free")

```


## Results

```{r plotPosterior}

plot(brm2)

brm2 %<>% recover_types(qdt_mass_act)

(epreds <- qdt_mass_act %>%
    data_grid(tilled) %>%
    add_epred_draws(brm2)) %>%
  ggplot(aes(x = .epred, y = tilled)) +
  stat_halfeye() +
  stat_pointinterval(.width = c(.66, .95)) +
  xlab("Seed productivity (g/m2)") +
  ylab("Tilled (Y/N)")


(Till_fx <- brm2 %>%
  spread_draws(b_tilledY) %>%
  ggplot(aes(x = b_tilledY)) +
  stat_halfeye() +
  geom_vline(xintercept = 0, linetype = "dashed")) 

```


The posterior median difference is `median(brm2 %>% spread_draws(b_tilledY) %>% pluck("b_tilledY"))` g/m2 with a standard error of `sd(brm2 %>% spread_draws(b_tilledY) %>% pluck("b_tilledY"))` g/m2.



