---
title: "Compare subunits across years by species"
author: "Rebeca Becdach"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

# CREATE A GRAPH WITH CI's FOR YIELD IN EACH SUBUNIT EACH YEAR BY SPECIES 
# AND REFUGE

## CI graphs

Create graphs for showing estimated means and CIs for subunits. One graph is
generated for each refuge and each species included in the survey. Each graph 
shows estimates from all years WHAP has been implemented. Graphs include all 
subunits surveyed in WHAP, including subunits that were not surveyed in one of 
the years on the graph.

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)
library(rmarkdown)
library(tidyverse)
library(knitr)
library(kableExtra)
library(purrr)
library(data.table)
library(janitor)

```


```{r prepData}

st_files <- list.files(path = "./outputTables/.",
                       pattern = "Swamp_TimothySppTbl_g_m2_\\d{4}\\.csv",
                       full.names = T)
wg_files <- list.files(path = "./outputTables/.",
                       pattern = "WatergrassSppTbl_g_m2_\\d{4}\\.csv",
                       full.names = T)
sw_files <- list.files(path = "./outputTables/.",
                       pattern = "SmartweedSppTbl_g_m2_\\d{4}\\.csv",
                       full.names = T)

whap_data <- mapply(c, st_files, wg_files, sw_files, SIMPLIFY = FALSE) %>%
  flatten() %>%
  set_names() %>%
  map_dfr(
    ~ read_csv(.x, col_types = cols(), col_names = TRUE),
    .id = "file_name"
  ) %>%
  mutate(species = str_extract(file_name, "Swamp_Timothy|Watergrass|Smartweed"),
         year = str_extract(file_name, "\\d{4}")) %>%
  separate(Unit, c("LIT", "unit", "subunit"), sep = "_") %>%
  # line below is a temp fix; SLW subunit names have repeats and are uninformative
  mutate(subunit = ifelse(LIT == "SLW", unit, subunit)) %>%
  clean_names() 

```


```{r viabilityClasses}

viaClass <- read.csv("WHAP_Viability.csv") %>%
  pivot_longer(cols = c(Poor, Fair, Good, Very_Good),
               names_to = "class",
               values_to = "lower_bound") %>%
  pivot_wider(names_from = LIT,
              values_from = lower_bound) %>%
  mutate(class = fct_reorder(class, KRN, mean),
         class_color = case_when(
           startsWith(as.character(class), "Poor") ~ "red",
           startsWith(as.character(class), "Fair") ~ "yellow",
           startsWith(as.character(class), "Good") ~ "chartreuse",
           startsWith(as.character(class), "Very_Good") ~ "chartreuse3"),
         class = factor(class, ordered = TRUE))

```


```{r CIPlot}

plot_years <- function(spp = c("total"),
                       data = whap_data) {
  
  for (i in unique(data$lit)) {
    
    data_i <- data %>% dplyr::filter(lit == i)
    
    for (j in unique(data_i$species)) {
      
      data_ij <- data_i %>% dplyr::filter(species == j)
      
      viaClass_j <- viaClass %>% dplyr::filter(Species == j)
      
      year_plot <-
        ggplot(data = data_ij %>% arrange(mass_g_m2),
               aes(x = subunit,
                   y = mass_g_m2,
                   color = year)) +
        scale_color_manual(values = c("2019" = "deepskyblue",
                                      "2020" = "dodgerblue3",
                                      "2021" = "midnightblue")) +
        labs(color = "year") +
        geom_rect(data = viaClass_j,
                  inherit.aes = FALSE,
                  aes(NULL,
                      NULL,
                      xmin = 0,
                      xmax = Inf,
                      fill = class),
                  ymin = unlist(viaClass_j[, i]),
                  ymax = unlist(c(viaClass_j[-1, i], Inf))) +
        scale_fill_manual(values = alpha(c("red",
                                           "yellow",
                                           "chartreuse",
                                           "chartreuse3"),
                                         0.2)) +
        geom_point(size = 3) +
        xlab('Subunit\n') +
        ylab(paste("\n", j, "seed head mass index (g/m^2)")) +
        geom_errorbar(aes(ymin = ifelse(ci90lwr_g_m2 < 0, 0, ci90lwr_g_m2),
                                   ymax = ci90upr_g_m2),
                      width = 0.3,
                      cex = 0.75) +
        theme(plot.title = element_text(size = 16, face = "bold"),
              axis.ticks.y = element_blank(),
              axis.text.x = element_text(face = "bold"),
              axis.title = element_text(size = 12,
                                        face = "bold"),
              strip.text.y = element_text(hjust = 0,
                                          vjust = 1,
                                          angle = 180,
                                          face = "bold")) +
        coord_flip()

    print(year_plot)
    }
  }
}
plot_years()

```


# CREATE A GRAPH WITH CI's FOR YIELD IN EACH SUBUNIT EACH YEAR BY SPECIES 
# AND REFUGE (ADD TO NEW FILE?)

## Seed head mass index bar charts

Create bar charts showing seed production by species and year for each refuge. 

```{r barchart}

# load in files
kgLIT_2021 <- read.csv("WHAP2021-22/ScriptInput2021/mass_kg_LIT_2021.txt")

kgLIT_2020 <- read.csv("WHAP2020-21/ScriptInput2020/mass_kg_LIT_2020.txt")

kgLIT_2019 <- read.csv("MainScriptDataOutput/mass_kg_LIT_2019.txt")


# prep data
gm2_LIT <- bind_rows(kgLIT_2019, kgLIT_2020, kgLIT_2021) %>%
  mutate(year = factor(year),
         st_g.m2 = st_kg.ha / 10,
         wg_g.m2 = wg_kg.ha / 10,
         sw_g.m2 = sw_kg.ha / 10) %>%
  dplyr::select(LIT, st_g.m2, wg_g.m2, sw_g.m2, year) %>%
  pivot_longer(cols = ends_with("g.m2"),
               names_to = "species",
               values_to = "g.m2",
               values_drop_na = TRUE)

gm2_LIT$species[gm2_LIT$species == "st_g.m2"] <- "swamp timothy"

gm2_LIT$species[gm2_LIT$species == "wg_g.m2"] <- "watergrass"

gm2_LIT$species[gm2_LIT$species == "sw_g.m2"] <- "smartweed"

# gm2_LIT <- gm2_LIT[-c(14), ]


# plot the bar charts
plot_gm2_LIT <- function(data = gm2_LIT) {
  
  year_bar_plots <- list()
  
  for (i in unique(data$LIT)) {
    
    data_i <- data %>% dplyr::filter(LIT == i)
    
    year_bar_plots[[i]] <-
                 ggplot(data = data_i,
                        aes(x = species,
                            y = g.m2,
                            fill = year)) +
                 # ylim(0, 250) + #set the same y-axis range for KRN and PIX
                 geom_bar(stat="identity", position=position_dodge()) +
                 scale_fill_manual(values = c("2019" = "deepskyblue",
                                              "2020" = "blue",
                                              "2021" = "darkblue")) +
                 xlab('\nspecies') +
                 ylab("seed head mass index (g/m2)\n") +
                 theme_minimal()
                
    print(year_bar_plots[[i]])
  }
  return(year_bar_plots)
}

plot_gm2_LIT()


```

