---
title: "d2m Model development and evaluation. SOP 5"
author: "Emilio A. Laca"
date: "11/13/2021"
editor_options: 
  chunk_output_type: console
always_allow_html: true
---


```{r setup, include=FALSE, echo=FALSE, message=FALSE}

knitr::opts_chunk$set(echo = TRUE)
library(kableExtra)
library(tidyverse)
library(car)
library(magrittr)
library(readxl)
library(emmeans)
library(lme4)
# library(lmerTest)
library(rsample)
library(recipes)
library(tidymodels)
library(LMERConvenienceFunctions)
library(multilevelmod)
library(knitr)
library(ggpubr)
library(caret)
library(furrr)
library(parallel)
library(doParallel)
library(parallelly)
library(tictoc)

```

# Introduction

# Data

```{r data, echo=FALSE, warning=FALSE, message=FALSE}

st_2019 <- readRDS("st_all_noO.rds") %>%
  rename_with(~gsub(".", "_", .x, fixed = TRUE)) %>%
  mutate(year = 2019,
         spp = "SwampTimothy",
         f2t_length_mm = NA,
         length_mm = avg_length_mm,
         width_mm = NA,
         mass_mg = mass_per_sh_mg,
         LIT = substr(id, 1, 3)) %>%
  dplyr::select(LIT,
                year,
                spp,
                f2t_length_mm,
                length_mm,
                width_mm,
                mass_mg)

wg_2019 <- readRDS("wgdat.rds") %>%
  rename_with(~gsub(".", "_", .x, fixed = TRUE)) %>%
  mutate(year = 2019,
         spp = "Watergrass",
         f2t_length_mm = F2T_cm * 10,
         length_mm = sh_len_cm * 10,
         width_mm = NA,
         mass_mg = mass_per_sh_mg,
         LIT = refuge) %>%
  dplyr::select(LIT,
                year,
                spp,
                f2t_length_mm,
                length_mm,
                width_mm,
                mass_mg)

mdc_2021 <- readRDS("sh_dim_MDC2021.rds") %>%
  mutate(year = 2021) %>%
  dplyr::select(LIT,
                year,
                spp,
                f2t_length_mm,
                length_mm,
                width_mm,
                mass_mg)

d2m_data <- bind_rows(st_2019, wg_2019, mdc_2021)

unique(d2m_data$LIT)

# Fix LIT codes
LitKey <- c(CLS = "CLS",
            KRN = "KRN",
            PXL = "PIX",
            SAC = "SAC",
            STL = "SLW",
            PIX = "PIX",
            SL_ = "SLW",
            Colusa = "CLS",
            Sacramento = "SAC",
            StoneLakes = "SLW",
            MDC = "MDC")

d2m_data$LIT <- dplyr::recode(d2m_data$LIT, !!!LitKey)

# Create a grouping variable for LIT and year
d2m_data %<>% mutate(LIT_yr = paste(LIT, year, sep = "_"),
                     log_mass_mg = log(mass_mg),
                     log_length_mm = log(length_mm),
                     log_f2t_length_mm = log(f2t_length_mm + 120))

wg_d2m <- d2m_data %>%
  dplyr::filter(spp == "Watergrass") %>%
  mutate(emerged = f2t_length_mm / length_mm,
         log_emerged = log(emerged + 0.5)) %>%
  arrange(LIT, year) %>%
  dplyr::filter(!is.nan(log_emerged))

st_d2m <- d2m_data %>%
  dplyr::select(-c(f2t_length_mm, log_f2t_length_mm)) %>%
  dplyr::filter(spp == "SwampTimothy") %>%
  arrange(LIT, year)

sw_d2m <- d2m_data %>%
  dplyr::filter(spp == "Smartweed") %>%
  dplyr::select(-c(f2t_length_mm, log_f2t_length_mm)) %>%
  mutate(vol_mm3 = pi * (width_mm/2)^2 * length_mm)


```

# Model development

## Models for Swamp Timothy

```{r st_plotData, echo=FALSE, warning=FALSE, message=FALSE}

scatterplot(mass_mg ~ length_mm | LIT_yr,
            smooth = FALSE,
#            log = "xy",
            data = st_d2m)

```


```{r defineInvBocCox}

# Define function to invert the Box-Cox transformation
invBoxCox <- function(x, lambda) {
  if (lambda == 0) exp(x) else (lambda * x + 1) ^ (1 / lambda)}
```

```{r st_modelRandomGroups, echo=FALSE, warning=FALSE, message=FALSE}

# Get optimal lambda
st_lamb1 <- powerTransform(lmer(mass_mg ~ log_length_mm +
                                  (log_length_mm|LIT_yr),
                                data = st_d2m)) %>%
  coef(round = TRUE) %>%
  unname()

# Lambda is 0, so a log transformation is used.

st_d2m_m3 <- lmer(bcPower(mass_mg, lambda = st_lamb1) ~
                    log_length_mm +
                    (log_length_mm|LIT_yr),
                  data = st_d2m)

summary(st_d2m_m3)

ranef(st_d2m_m3) %>%
  as.data.frame() %>%
  dplyr::select(term,
                grp,
                condval) %>%
  pivot_wider(names_from = term,
              values_from = condval) %>%
  rename(Group = grp,
         Intercept = `(Intercept)`,
         Slope = log_length_mm) %>%
  kable(digits = c(0, 3, 3), caption = "Random effects of group on the slope and intercept of the relationship between log mass and log length of Swamp Timothy seed heads.") %>%
  kable_styling(full_width = FALSE)



```

```{r st_plotRanef, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE, eval=FALSE}

rnf_st_d2m_m3 <- ranef(st_d2m_m3) %>%
  as.data.frame() %>%
  pivot_wider(id_cols = "grp",
              values_from = c(condval, condsd),
              names_from = "term") %>%
  rename(cond_Incpt_Ranef = `condval_(Intercept)`,
         cond_slope_Ranef = `condval_log_length_mm`,
         cond_Incpt_sd = `condsd_(Intercept)`,
         cond_slope_sd = `condsd_log_length_mm`)

ggplot(data = rnf_st_d2m_m3,
       aes(x = cond_Incpt_Ranef,
           y = cond_slope_Ranef)) +
  geom_point(color = "blue", size = 3) + 
  geom_errorbar(aes(ymin = cond_slope_Ranef - cond_slope_sd,
                    ymax = cond_slope_Ranef + cond_slope_sd),
                color = "blue") +
  geom_errorbarh(aes(xmin = cond_Incpt_Ranef - cond_Incpt_sd,
                     xmax = cond_Incpt_Ranef + cond_Incpt_sd),
                 color = "orange") +
  geom_label(aes(label = grp), hjust = 0.5, vjust = 0.3) +
  xlab("Random effect on intercept") +
  ylab("Random effect on slope")

```

```{r st_plotAllData, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE, eval=FALSE}

ggplot(data = st_d2m,
       aes(x = predict(st_d2m_m3),
           y = log_mass_mg,
           color = LIT_yr,
           groups = LIT_yr)) +
  geom_point() +
  geom_smooth(method = "lm",
              se = TRUE) +
  geom_abline() +
  facet_wrap(~LIT_yr,
             ncol = 3) +
  theme(legend.position = "none")

```

```{r st_bootMer, echo=FALSE, warning=FALSE, message=FALSE}

St_CoefVar <- st_d2m %>%
  group_by(LIT_yr) %>%
  summarise(log_length_mm = log(mean(length_mm, na.rm = TRUE)),
            .groups = "drop")

myPred_st <- function(.) {
  invBoxCox(predict(.,
                    newdata = St_CoefVar,
                    re.form = NULL),
            lambda = st_lamb1)
}

# Based on
# https://cran.r-project.org/web/packages/merTools/vignettes/Using_predictInterval.html
sumBoot <- function(merBoot_t) {
  return(
    data.frame(
      mass_median = apply(merBoot_t,
                          2,
                          function(x) as.numeric(quantile(x,
                                                          probs = 0.50,
                                                          na.rm = TRUE))),
      mass_mean = apply(merBoot_t,
                        2,
                        function(x) as.numeric(mean(x,
                                                    na.rm = TRUE))),
      lwr = apply(merBoot_t,
                  2, function(x) as.numeric(quantile(x,
                                                     probs = .05,
                                                     na.rm = TRUE))),
      upr = apply(merBoot_t,
                  2, function(x) as.numeric(quantile(x,
                                                     probs = .95,
                                                     na.rm = TRUE))),
      se = apply(merBoot_t,
                 2, function(x) as.numeric(sd(x,
                                              na.rm = TRUE)))
    )
  )
}

# Use bootMer to get estimates of prediction intervals and se's
boot1_st <- lme4::bootMer(st_d2m_m3,
                       myPred_st,
                       nsim = 1500,
                       parallel = "multicore",
                       ncpus = 8,
                       use.u = TRUE,
                       type = "parametric")


St_CoefVar %<>%
  bind_cols(sumBoot(boot1_st$t)) %>%
  mutate(cv2 = sqrt(se^2 / (mass_mean^2 + se^2)))

rownames(St_CoefVar) <- NULL

my_hist <- function(x, y) hist(x,
                            breaks = 25,
                            xlim = c(0, 60),
                            freq = FALSE,
                            col = "skyblue",
                            xlab = "Prediction (mg)",
                            main = y)

colnames(boot1_st$t) <- St_CoefVar$LIT_yr


opar <- par(mfrow = c(3,2)); x <- map2(.x = as.data.frame(boot1_st$t),
                                       .y = colnames(boot1_st$t),
                                       ~my_hist(.x, .y)); par(opar)

St_CoefVar %<>%
  dplyr::select(LIT_yr,
                mass_median,
                lwr,
                mass_mean,
                upr,
                se,
                cv2) %>%
    mutate(MOE = (upr - lwr) / (upr + lwr),
           n = as.numeric(with(st_d2m,  table(LIT_yr))))

St_CoefVar %>%
  kable(caption = "Estimated seed head mass (mg), coefficient of variation (cv2), margin of error (MOE) and 90% CI for swamp timothy seedheads of average dimensions.",
        digits = c(0, 1, 1, 1, 1, 1, 2, 2, 0)) %>%
  kable_styling(full_width = FALSE)

```

## Models for Watergrass

```{r wg_plotData, echo=FALSE, warning=FALSE, message=FALSE}

pairs(wg_d2m[ c("length_mm", "f2t_length_mm", "emerged")])

scatterplot(mass_mg ~ length_mm | LIT_yr,
            smooth = FALSE,
            log = "xy",
            data = wg_d2m)

scatterplot(mass_mg ~ log_f2t_length_mm | LIT_yr,
            smooth = FALSE,
            log = "y",
            data = wg_d2m)

scatterplot(mass_mg ~ log_emerged | LIT_yr,
            smooth = FALSE,
            log = "y",
            data = wg_d2m)


```

```{r wg_modelRandomGroups, echo=TRUE, warning=FALSE, message=FALSE}

wg_frml1 <- mass_mg ~
  log_length_mm +
  log_emerged +
  log_length_mm:log_emerged +
  (1 | LIT_yr) +
  (0 + log_length_mm | LIT_yr) +
  (0 + log_emerged | LIT_yr) + 
  (0 + log_length_mm:log_emerged | LIT_yr)

wg_lamb = powerTransform(
  lmer(wg_frml1,
       data = wg_d2m)) %>%
  coef(round = TRUE) %>%
  unname()

# wg_d2m %<>%
#   mutate(bc_mass = bcPower(mass_mg, lambda = wg_lamb))
# 
wg_d2m_m2 <- lmer(update(wg_frml1, bcPower(mass_mg, lambda = wg_lamb) ~ .),
                  data = wg_d2m)

summary(wg_d2m_m2)

wg_d2m_m3 <- update(wg_d2m_m2, . ~ . - (0 + log_length_mm:log_emerged | LIT_yr))

# wg_d2m_m2 is better than wg_d2m_m3. Try to remove another random effect.
anova(wg_d2m_m2, wg_d2m_m3)

wg_d2m_m4 <- update(wg_d2m_m2, . ~ . -(0 + log_length_mm | LIT_yr))

anova(wg_d2m_m2, wg_d2m_m4) # wg_d2m_m4 is better

summary(wg_d2m_m4)

# Re fit wg_d2m_m4 with a new lambda

wg_lamb = powerTransform(
  lmer(update(formula(wg_d2m_m4), mass_mg ~ .),
       data = wg_d2m)) %>%
  coef(round = TRUE) %>%
  unname()

wg_d2m_m4 <- lmer(formula(wg_d2m_m4), data = wg_d2m)

summary(wg_d2m_m4)

wg_d2m_m5 <- update(wg_d2m_m4, . ~ . -(0 + log_length_mm:log_emerged | LIT_yr))

anova(wg_d2m_m5, wg_d2m_m2)

summary(wg_d2m_m5)

print("Pseudo R-squared for model wg_d2m_m4")
MuMIn::r.squaredGLMM(wg_d2m_m4)

print("Pseudo R-squared for model wg_d2m_m5")
MuMIn::r.squaredGLMM(wg_d2m_m5)

```

```{r wg_bootMer, echo=FALSE, warning=FALSE, message=FALSE}

Wg_CoefVar <- wg_d2m %>%
  group_by(LIT_yr) %>%
  summarise(log_length_mm = log(mean(length_mm, na.rm = TRUE)),
            log_f2t_length_mm = log(mean(f2t_length_mm, na.rm = TRUE)),
            log_emerged = log(mean(emerged, na.rm = TRUE)),
            .groups = "drop")

myPred <- function(.) {
  invBoxCox(predict(.,
              newdata = Wg_CoefVar,
              re.form = NULL),
            lambda = wg_lamb)
}

# Use bootMer to get estimates of prediction intervals and se's

boot1_wg <- lme4::bootMer(wg_d2m_m4,
                       myPred,
                       nsim = 1500,
                       parallel = "multicore",
                       ncpus = 8,
                       use.u = TRUE,
                       type = "parametric")


Wg_CoefVar %<>%
  bind_cols(sumBoot(boot1_wg$t)) %>%
  mutate(cv2 = sqrt(se^2 / (mass_mean^2 + se^2)))

rownames(Wg_CoefVar) <- NULL

my_hist <- function(x, y) hist(x,
                            breaks = 25,
                            xlim = c(0, max(x)),
                            freq = FALSE,
                            col = "skyblue",
                            xlab = "Prediction (mg)",
                            main = y)

colnames(boot1_wg$t) <- Wg_CoefVar$LIT_yr


opar <- par(mfrow = c(2,2)); x <- map2(.x = as.data.frame(boot1_wg$t),
                                  .y = colnames(boot1_wg$t),
                                  ~my_hist(.x, .y)); par(opar)

rm(x)

Wg_CoefVar %<>%
  dplyr::select(LIT_yr,
                mass_median,
                lwr,
                mass_mean,
                upr,
                se,
                cv2) %>%
  mutate(MOE = (upr - lwr) / (upr + lwr),
           n = as.numeric(with(wg_d2m,  table(LIT_yr))))

Wg_CoefVar %>%
  kable(caption = "Estimated seed head mass (mg), coefficient of variation (cv2), margin of error (MOE) and 90% CI for watergrass seedheads of average dimensions. Based on bootMer",
        digits = c(0, 0, 0, 0, 0, 1, 2, 2, 0)) %>%
  kable_styling(full_width = FALSE)

```

```{r rsq}

MuMIn::r.squaredGLMM(wg_d2m_m4)

MuMIn::r.squaredGLMM(wg_d2m_m5)

```

```{r tidyBootstrapWg, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE, eval=FALSE}

# Not run.
# This chunk simply corroborates the results of bootMer using tidy bootstrapping.
# It assumes that the transformation is constant (constant lambda)
# Results were corroborated, so this additional code is not run or presented.

Wg_CoefVar <- wg_d2m %>%
  group_by(LIT_yr) %>%
  summarise(log_length_mm = log(mean(length_mm, na.rm = TRUE)),
            log_f2t_length_mm = log(mean(f2t_length_mm, na.rm = TRUE)),
            log_emerged = log(mean(emerged, na.rm = TRUE)),
            .groups = "drop")

# Use tidymodels to get bootstraps of predicted mass per seed head.
wg_boots <- bootstraps(wg_d2m,
                       times = 1500,
                       strata = LIT_yr,
                       apparent = TRUE)

wg_pred_boots <- function(split) {
  # wg_lamb = powerTransform(lmer(update(formula(wg_d2m_m4), # for variable lambda
  #                                      mass_mg ~ .),
  #                               data =  analysis(split))) %>%
  #   coef(round = TRUE) %>%
  #   unname()
  
  wg_m4_boot = lmer(update(formula(wg_d2m_m4),
                           bcPower(mass_mg, lambda = wg_lamb) ~ .),
                    data = analysis(split))
  
  mass_mg = predict(wg_m4_boot,
                    newdata = Wg_CoefVar,
                    re.form = NULL) %>%
    invBoxCox(lambda = wg_lamb) %>%
    as_tibble_row() %>%
    setNames(nm = Wg_CoefVar$LIT_yr)
  return(mass_mg)
}

tic()
wg_boots %<>% 
  mutate(pred = mclapply(X = wg_boots$splits,
                         FUN = function(x) wg_pred_boots(x)),
                         mc.cores = 8) %>%
  unnest(pred)
toc()

Wg_CoefVar %<>%
  bind_cols(wg_boots %>%
              dplyr::select(Wg_CoefVar$LIT_yr) %>%
              sumBoot()) %>%
  mutate(cv2 = sqrt(se^2 / (mass_mean^2 + se^2)))

rownames(Wg_CoefVar) <- NULL

my_hist <- function(x, group) hist(x,
                            breaks = 30,
                            xlim = c(0, max(x)),
                            freq = FALSE,
                            col = "skyblue",
                            xlab = "Prediction (mg)",
                            main = group)

opar <- par(mfrow = c(2,2)); xx <- map2(.x = wg_boots %>%
                                          dplyr::select(Wg_CoefVar$LIT_yr),
                                  .y = Wg_CoefVar$LIT_yr,
                                  ~my_hist(.x, .y)); par(opar)

rm(xx)

Wg_CoefVar %<>%
  dplyr::select(LIT_yr,
                mass_median,
                lwr,
                mass_mean,
                upr,
                se,
                cv2) %>%
  mutate(MOE = (upr - lwr) / (upr + lwr),
           n = as.numeric(with(wg_d2m,  table(LIT_yr))))

Wg_CoefVar %>%
  kable(caption = "Estimated seed head mass (mg), coefficient of variation (cv2), margin of error (MOE) and 90% CI for watergrass seedheads of average dimensions. Based on tidybootstraps.",
        digits = c(0, 0, 0, 0, 0, 1, 2, 2)) %>%
  kable_styling(full_width = FALSE)


```

## Models for Smartweed

```{r sw_exploreData, echo=FALSE, warning=FALSE, message=FALSE}

sw_d2m %>% dplyr::select(mass_mg, length_mm, width_mm, vol_mm3) %>% pairs()

scatterplot(mass_mg ~ vol_mm3, sw_d2m, log = "xy") # Log transf. is promising

sw_d2m[which.max(sw_d2m$vol_mm3), ] # extreme volume

sw_d2m %>% dplyr::select(length_mm, width_mm) %>% plot()
abline(0, 1) # One seed head has width 8 and length 2; nonsense. remove

which(sw_d2m$length_mm < 4 & sw_d2m$width_mm == 8)

sw_d2m %<>% dplyr::filter(
  not(vol_mm3 == max(vol_mm3) | (sw_d2m$length_mm < 4 & sw_d2m$width_mm == 8))
) # to exclude oddballs

```

#### Volume as single predictor

```{r sw_linearVolume, echo=FALSE, warning=FALSE, message=FALSE}

sw_m1 <- lm(mass_mg ~ vol_mm3, data = sw_d2m)

summary(sw_m1)

opar <- par(mfrow = c(2,2)); plot(sw_m1); par(opar)

```

```{r sw_transformedVolume, echo=TRUE, warning=FALSE, message=FALSE}

# Get optimal lambda
sw_lamb1 <- powerTransform(lm(mass_mg ~ vol_mm3,
                              data = sw_d2m)) %>%
  coef(round = TRUE) %>%
  unname()

# Fit model with transformed Y
sw_m2 <- lm(bcPower(mass_mg, lambda = sw_lamb1) ~ vol_mm3,
               data = sw_d2m)

opar <- par(mfrow = c(2,2)); plot(sw_m2); par(opar)

# Some nonlinearity remains, which increases the variance at the right side
# Try an additional transformation in volume

sw_lamb2 <- powerTransform(lm(mass_mg ~ log(vol_mm3),
                              data = sw_d2m)) %>%
  coef(round = TRUE) %>%
  unname()

sw_m3 <- lm(bcPower(mass_mg, lambda = sw_lamb2) ~ log(vol_mm3),
               data = sw_d2m)

opar <- par(mfrow = c(2,2)); plot(sw_m3); par(opar)

summary(sw_m3)

```

```{r sw_tidyBootstrapSwVol, echo=FALSE, warning=FALSE, message=FALSE}

Sw_CoefVar <- sw_d2m %>%
  group_by(LIT_yr) %>%
  summarise(log_length_mm = log(mean(length_mm, na.rm = TRUE)),
            vol_mm3 = mean(vol_mm3, na.rm = TRUE),
            .groups = "drop")

# Use tidymodels to get bootstraps of predicted mass per seedhead.
sw_boots <- bootstraps(sw_d2m,
                       times = 1500,
                       apparent = TRUE)

sw_pred_boots <- function(split) {
  sw_m3_boot = lm(bcPower(mass_mg, lambda = sw_lamb2) ~ log(vol_mm3),
                  data = analysis(split)) 
  mass_mg = predict(sw_m3_boot, newdata = Sw_CoefVar) %>%
    invBoxCox(lambda = sw_lamb2)
  return(mass_mg)
}

sw_boots %<>% 
  mutate(pred = map_dbl(splits, sw_pred_boots))

Sw_CoefVar %<>%
  bind_cols(sw_boots$pred %>%
              as.matrix() %>%
              sumBoot()) %>%
  mutate(cv2 = sqrt(se^2 / (mass_mean^2 + se^2)))

rownames(Sw_CoefVar) <- NULL

hist(sw_boots$pred,
     breaks = 25,
     xlim = c(5, max(sw_boots$pred)),
     freq = FALSE,
     col = "skyblue",
     xlab = "Prediction (mg)",
     main = "Predicted seed head mass. MDC_2021")

Sw_CoefVar %<>%
  dplyr::select(LIT_yr,
                mass_median,
                lwr,
                mass_mean,
                upr,
                se,
                cv2) %>%
  mutate(MOE = (upr - lwr) / (upr + lwr),
         n = length(sw_d2m$mass_mg))

Sw_CoefVar %>%
  kable(caption = "Estimated seed head mass (mg), coefficient of variation (cv2), margin of error (MOE) and 90% CI for Smartweed seedheads of average dimensions. Model uses log(vol_mm3) as single predictor.",
        digits = c(0, 0, 0, 0, 0, 1, 2, 2, 0)) %>%
  kable_styling(full_width = FALSE)

```

#### Multiple predictors

```{r sw_multiplePreds, echo=FALSE, warning=FALSE, message=FALSE}

sw_lamb3 <- powerTransform(lm(mass_mg ~ log(vol_mm3) + width_mm + length_mm,
                              data = sw_d2m)) %>%
  coef(round = TRUE) %>%
  unname()

sw_m4 <- lm(bcPower(mass_mg, lambda = sw_lamb3) ~ log(vol_mm3) +
              width_mm + length_mm,
            data = sw_d2m)

opar <- par(mfrow = c(2,2)); plot(sw_m4); par(opar)

summary(sw_m4)
anova(sw_m4)

```

```{r sw_tidyBootstrapSwMulti, echo=FALSE, warning=FALSE, message=FALSE}

Sw_CoefVar <- sw_d2m %>%
  group_by(LIT_yr) %>%
  summarise(length_mm = mean(length_mm, na.rm = TRUE),
            width_mm = mean(width_mm, na.rm = TRUE),
            vol_mm3 = mean(vol_mm3, na.rm = TRUE),
            .groups = "drop")

# Use tidymodels to get bootstraps of predicted mass per seedhead.
sw_boots <- bootstraps(sw_d2m,
                       times = 1500,
                       apparent = TRUE)

sw_pred_boots <- function(split) {
  sw_m4_boot = lm(bcPower(mass_mg, lambda = sw_lamb3) ~
                    log(vol_mm3) +
                    width_mm +
                    length_mm,
                  data = analysis(split)) 
  mass_mg = predict(sw_m4_boot, newdata = Sw_CoefVar) %>%
    invBoxCox(lambda = sw_lamb3)
  return(mass_mg)
}

sw_boots %<>% 
  mutate(pred = map_dbl(splits, sw_pred_boots))

Sw_CoefVar %<>%
  bind_cols(sw_boots$pred %>%
              as.matrix() %>%
              sumBoot()) %>%
  mutate(cv2 = sqrt(se^2 / (mass_mean^2 + se^2)))

rownames(Sw_CoefVar) <- NULL

hist(sw_boots$pred,
     breaks = 25,
     xlim = c(5, max(sw_boots$pred)),
     freq = FALSE,
     col = "skyblue",
     xlab = "Prediction (mg)",
     main = "Predicted seed head mass. MDC_2021")

Sw_CoefVar %<>%
  dplyr::select(LIT_yr,
                mass_median,
                lwr,
                mass_mean,
                upr,
                se,
                cv2) %>%
  mutate(MOE = (upr - lwr) / (upr + lwr),
         n = length(sw_d2m$mass_mg))

Sw_CoefVar %>%
  kable(caption = "Estimated seed head mass (mg), coefficient of variation (cv2), margin of error (MOE) and 90% CI for Smartweed seedheads of average dimensions. Model uses log(vol_mm3), length_mm and width_mm as predictors.",
        digits = c(0, 0, 0, 0, 0, 1, 2, 2, 0)) %>%
  kable_styling(full_width = FALSE)

```

#### Length as a single predictor

```{r sw_transformedLength, echo=FALSE, warning=FALSE, message=FALSE}

# Get optimal lambda
sw_lamb4 <- powerTransform(lm(mass_mg ~ length_mm,
                              data = sw_d2m)) %>%
  coef(round = TRUE) %>%
  unname()

# Fit model with transformed Y
sw_m5 <- lm(bcPower(mass_mg, lambda = sw_lamb4) ~ length_mm,
               data = sw_d2m)

opar <- par(mfrow = c(2,2)); plot(sw_m5); par(opar)

# Some nonlinearity remains, which increases the variance at the right side
# Try an log transformation of length

# Get new optimal lambda
sw_lamb5 <- powerTransform(lm(mass_mg ~ log_length_mm,
                              data = sw_d2m)) %>%
  coef(round = TRUE) %>%
  unname()

sw_m6 <- lm(bcPower(mass_mg, lambda = sw_lamb5) ~ log_length_mm,
               data = sw_d2m)

opar <- par(mfrow = c(2,2)); plot(sw_m6); par(opar)

summary(sw_m5)
anova(sw_m5)

```

```{r sw_tidyBootstrapSwLen, echo=FALSE, warning=FALSE, message=FALSE}

Sw_CoefVar <- sw_d2m %>%
  group_by(LIT_yr) %>%
  summarise(log_length_mm = log(mean(length_mm, na.rm = TRUE)),
            .groups = "drop")

# Use tidymodels to get bootstraps of predicted mass per seed head.
sw_boots <- bootstraps(sw_d2m,
                       times = 1500,
                       apparent = TRUE)

sw_pred_boots <- function(split) {
  sw_m6_boot = lm(bcPower(mass_mg, lambda = sw_lamb5) ~ log_length_mm,
                  data = analysis(split)) 
  mass_mg = predict(sw_m6_boot, newdata = Sw_CoefVar) %>%
    invBoxCox(lambda = sw_lamb5)
  return(mass_mg)
}

sw_boots %<>% 
  mutate(pred = map_dbl(splits, sw_pred_boots))

Sw_CoefVar %<>%
  bind_cols(sw_boots$pred %>%
              as.matrix() %>%
              sumBoot()) %>%
  mutate(cv2 = sqrt(se^2 / (mass_mean^2 + se^2)))

rownames(Sw_CoefVar) <- NULL

hist(sw_boots$pred,
     breaks = 25,
     xlim = c(5, max(sw_boots$pred)),
     freq = FALSE,
     col = "skyblue",
     xlab = "Prediction (mg)",
     main = "Predicted seed head mass. MDC_2021")

Sw_CoefVar %<>%
  dplyr::select(LIT_yr,
                mass_median,
                lwr,
                mass_mean,
                upr,
                se,
                cv2) %>%
  mutate(MOE = (upr - lwr) / (upr + lwr),
         n = length(sw_d2m$mass_mg))

Sw_CoefVar %>%
  kable(caption = "Estimated seed head mass (mg), coefficient of variation (cv2), margin of error (MOE) and 90% CI for Smartweed seedheads of average dimensions. Model uses log(length_mm) as single predictor.",
        digits = c(0, 0, 0, 0, 0, 1, 2, 2, 0)) %>%
  kable_styling(full_width = FALSE)

```

# Model Evaluation

### Function to evaluate alternatives

```{r evalFun, echo=FALSE, warning=FALSE, message=FALSE}
# Model option mld.opt has to be either "fixed" or anything else. Anything else
# results in the "mixed" model.
model_eval_fnc <- function(the_data = NULL) { # Start f() definition ******
  
  # Make vector of LIT_yr groups to loop over
  grps <- unique(the_data$LIT_yr)
  
  # Create data frames to receive output from validation (test)
  test_metrics <- tibble(grp_out = character(),
                         .metric =  character(),
                         .estimate = double())
  
  test_data4graphs <- tibble(LIT_yr = character(),
                             mass_mg =  double(),
                             .pred = double())
  
  # Determine what species is being analyzed
  spp <- the_data %>%
    pluck("spp") %>%
    unique()
  
  #Prepare data for manual initial split by LIT_yr
  the_data %<>%
    arrange(LIT_yr) %>%
    mutate(.row = row_number())
  
  # Define model
  lm_model <-
    linear_reg() %>%
    set_engine("lmer")
  
  
  # Define formulae
  frmls <- list(SwampTimothy = formula(st_d2m_m3),
                Watergrass = formula(wg_d2m_m4))
  
  # Define lambdas
  lambdas <- list(SwampTimothy = st_lamb1,
                Watergrass = wg_lamb)
  
  
  m_formula <- frmls[[spp]]
  
  # Start of cross-validation to assess option 1
  df_vfold <- vfold_cv(the_data,
                       v = 6,
                       repeats = 10,
                       strata = LIT_yr,
                       pool = 0.01)
  
  # Extract analysis and assessment data
  cv_out <- mutate(df_vfold,
                   df_cv_trn = map(splits,  analysis),
                   df_cv_tst = map(splits,  assessment)) %>%
    # Fit model to all folds
    mutate(model_fit = map(df_cv_trn,
                           ~fit(lm_model, 
                                m_formula, 
                                data = .x))) %>% 
    # Predict on assessment fold and save observed and predicted
    mutate(model_pred = map2(model_fit,
                             df_cv_tst,
                             ~invBoxCox(predict(.x,
                                                new_data = .y,
                                                type = "raw",
                                                opts = list(re.form = NULL)),
                                        lambda = lambdas[[spp]]))) %>% 
    mutate(reslt = map2(df_cv_tst,
                        model_pred,
                        ~data.frame(LIT_yr = .x$LIT_yr,
                                    mass_mg = .x$mass_mg,
                                    .pred = .y)))
  
  # Save test data and predictions for graphs
  cv_out4graphs <- cv_out %>%
    dplyr::select(id, df_cv_tst, model_pred) %>%
    unnest(cols = c(df_cv_tst, model_pred)) %>%
    arrange(id, LIT_yr)
  
  # Calculate and save metrics
  cv_metrics <- cv_out %>%
    dplyr::select(id, reslt) %>% 
    tidyr::unnest(reslt) %>% 
    group_by(LIT_yr) %>%
    # Get metrics to evaluate model
    metrics(truth = mass_mg, estimate = .pred) %>%
    dplyr::select(LIT_yr,
                  .metric,
                  .estimate) %>%
    pivot_wider(names_from = .metric,
                values_from = .estimate)
  
  # Start of validation to assess option 2
  for (grp in grps) { # Start loop over LIT_yr groups for validation
    
    # Split data leaving one group out at a time as testing data.
    indices <-
      list(analysis   = the_data$.row[the_data$LIT_yr != grp], 
           assessment = the_data$.row[the_data$LIT_yr == grp])
    
    df_split <- make_splits(indices, the_data %>% dplyr::select(-.row))
    df_train <- training(df_split)
    df_test  <- testing(df_split)
    
    # Fit model to all training data for prediction in testing data (grp)
    all_fit <- fit(lm_model, m_formula, data = df_train)
    
    lambda_val <- lambdas[[spp]]
    
    df_test %<>% 
      mutate(model_pred = predict(all_fit,
                                  new_data = df_test,
                                  type = "raw",
                                  opts = list(allow.new.levels = TRUE))) %>%
      mutate(model_pred = invBoxCox(model_pred, lambda = lambda_val))
    
    test_metrics <- metrics(df_test, truth = mass_mg, estimate = model_pred) %>%
      mutate(grp_out = grp) %>%
      dplyr::select(grp_out, .metric, .estimate) %>%
      bind_rows(test_metrics, .)
    
    test_data4graphs <- df_test %>%
      bind_rows(test_data4graphs, .)
    
  } # End loop over LIT_yr groups *****
  
  test_metrics %<>%
    pivot_wider(names_from = .metric,
                values_from = .estimate)
  
  return(list(species = spp,
              cv_metrics = cv_metrics,
              cv_out4graphs = cv_out4graphs,
              test_metrics = test_metrics,
              test_data4graphs = test_data4graphs))
  
} # End function definition ******

```

### Function to summarize results of each evaluation

```{r summarizePlotFun, echo=FALSE, warning=FALSE, message=FALSE}

summarize_plot_fnc <- function(rslt_list = NULL) {# Start summarize_plot function definition
  
  # Checks for correctness of list, etc. TBA

  # Cross validation summary
  cross_val_metrics <- rslt_list$cv_metrics %>%
    dplyr::select(-LIT_yr) %>%
    summarize_all(mean) %>%
    mutate(LIT_yr = "Average") %>%
    bind_rows(rslt_list$cv_metrics, .)
  
  Tbl_cv_cap <- paste("Table of cross validation results for ", rslt_list$species, ". These results represent how well the model performs when all groups have data for d2m.", sep = "")
  
  # Cross validation table
  tbl_cv <-  knitr::kable(cross_val_metrics,
                          digits = c(0, 1, 2, 1),
                          caption = Tbl_cv_cap) %>%
    kable_styling(full_width = FALSE)
  
  # Cross validation figure
  fig_cv <-  ggplot(data = rslt_list$cv_out4graphs,
                    aes(x = model_pred,
                        y = mass_mg,
                        color = LIT_yr,
                        groups = id)) +
    geom_smooth(method = "lm",
                se = TRUE,
                size = 0.01,
                alpha = 0.20) +
    geom_smooth(method = "lm",
                se = FALSE) +
    geom_point() +
    geom_abline() +
    facet_wrap(~LIT_yr,
               ncol = 2) +
    theme(legend.position = "none") + 
    annotate("text",
             x = 0.90 * max(rslt_list$cv_out4graphs$model_pred, na.rm = TRUE),
             y = 0.15 * max(rslt_list$cv_out4graphs$mass_mg, na.rm = TRUE),
             label = rslt_list$species)
  
  
  # Validation summary
  val_metrics <- rslt_list$test_metrics %>%
    dplyr::select(-grp_out) %>%
    summarize_all(mean) %>%
    mutate(grp_out = "Average") %>%
    bind_rows(rslt_list$test_metrics, .)
  
  Tbl_v_cap <- paste("Table of final validation results for ", rslt_list$species, " when data for each refuge is kept out from model creation and used for final testing. This test provides a realistic estimate of the error expected when using data from an observed set of refuge-year combinations to predict seed head mass in a new refuge-year.", sep = "")
  
  # Validation table
  tbl_v <-  knitr::kable(val_metrics,
                         digits = c(0, 1, 2, 1),
                         caption = Tbl_v_cap) %>%
    kable_styling(full_width = FALSE)
  
  # Validation metrics figure
  fig_v <- ggplot(data = rslt_list$test_data4graphs,
                  aes(x = model_pred,
                      y = mass_mg,
                      color = LIT_yr,
                      groups = LIT_yr)) +
    geom_point() +
    geom_smooth(method = "lm",
                se = TRUE) +
    geom_abline() +
    facet_wrap(~LIT_yr,
               ncol = 2) +
    theme(legend.position = "none") +
    annotate("text",
             x = 0.90 * max(rslt_list$test_data4graphs$model_pred, na.rm = TRUE),
             y = 0.15 * max(rslt_list$test_data4graphs$mass_mg, na.rm = TRUE),
             label = rslt_list$species)
  
  return(list(cv_table = tbl_cv,
              cv_fig = fig_cv,
              v_table = tbl_v,
              v_fig = fig_v))
  
}

```

## Evaluation of prediction alternatives

```{r evalOptions, echo=FALSE, warning=FALSE, message=FALSE}

# Swam Timothy
st_model_eval <- model_eval_fnc(the_data = st_d2m)
st_rslts <- summarize_plot_fnc(rslt_list = st_model_eval)

# Watergrass
wg_model_eval <- model_eval_fnc(the_data = wg_d2m)
wg_rslts <- summarize_plot_fnc(rslt_list = wg_model_eval)

```

### Swamp Timothy graphical results

```{r displayGraphsSwampTimothy, echo=FALSE, warning=FALSE, message=FALSE}

print("SWAMP TIMOTHY PREDICTIONS FOR SAMPLED REFUGES")
st_rslts$cv_fig

print("SWAMP TIMOTHY PREDICTIONS FOR NEW REFUGES")
st_rslts$v_fig

```

### Swamp Timothy metrics

```{r displayTablesSwampTimothy, echo=FALSE, warning=FALSE, message=FALSE}

st_rslts$cv_table %>%
  kable_styling(full_width = FALSE,
                position = "float_left")

st_rslts$v_table %>%
  kable_styling(full_width = FALSE,
                position = "left")


```

### Watergrass graphical results

```{r displayGraphsWatergrass, echo=FALSE, warning=FALSE, message=FALSE}

print("WATERGRASS PREDICTIONS FOR SAMPLED REFUGES")
wg_rslts$cv_fig

print("WATERGRASS PREDICTIONS FOR NEW REFUGES")
wg_rslts$v_fig

```

### Watergrass metrics

```{r displayTablesWatergrass, echo=FALSE, warning=FALSE, message=FALSE}

wg_rslts$cv_table %>%
  kable_styling(full_width = FALSE,
                position = "float_left")

wg_rslts$v_table %>%
  kable_styling(full_width = FALSE,
                position = "left")


```
