---
title: "WHAP2 User Interface"
author: "Emilio A. Laca"
date: "`r format(Sys.time(), '%a %d %b %Y %X')`"
output: html_document
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

*Original date: 3 May 2022*

This user interface should have sections that parallel the scope of work. The user should have options to select years, refuges, subunits, etc. to be included in the report.

1. Data intake and validation.
2. Calculations for any number of species and refuges.
3. Analysis of temporal trends in any subset of subunits.
4. Comparison of groups of subunits that had different management.

Data intake and validation is covered in WHAP2_DataChecks.Rmd, which should probably be transformed into .R, maybe as a function.  

Calculations are performed through:

d2m_modeling.R
SimMassBySuLITFun.R
SimMassPerAreaFun.R
SimPropAreaFun.R
WHAP2_DataPrep.R
WHAP2_sim_stats.Rmd

The directory organization that has to be preserved for the software to work is:

WHAP2_SimBasedEstimates (folder)
.
|____inputFiles (folder)
|  |____qdt_2021.rds
|  |____qdt_2020.rds
|  |____qdt_2019.rds
|  |____vpcp2019.rds
|  |____vpcp2020.rds
|  |____vpcp2021.rds
|
|____SimPropAreaFun.R
|
|____outputFiles (folder)
|  |____shmi_2020.rds
|  |____shmi_2021.rds
|  |____shmi_2019.rds
|
|____d2mFiles (folder)
|  |____files_already_processed.rds
|  |____d2m_models_list.rds
|  |____d2m_data.rds
|
|____WHAP2_DataPrep.R
|____WHAP2_sim_stats.Rmd
|____d2m_modeling.R
|____SimMassPerAreaFun.R
|____WHAP2_User_Interface.Rmd
|
|____ProcessedFieldData (folder)
|  |____WHAP_2021_VantagePolygons_20210831.csv
|  |____WHAP_2021_Quadrats_20210923.csv
|  |____WHAP_2021_CirclePlots_20210901.csv
|  |____WHAP_2021_ManagementUnits_20210902.csv
|
|____SimMassBySuLITFun.R
|____WHAP2_DataChecks.Rmd



## Inputs and Outputs

You can embed Shiny inputs and outputs in your document. Outputs are automatically updated whenever inputs change.  This demonstrates how a standard R plot can be made interactive by wrapping it in the Shiny `renderPlot` function. The `selectInput` and `sliderInput` functions create the input widgets used to drive the plot.

```{r eruptions, echo=FALSE}
inputPanel(
  selectInput("n_breaks", label = "Number of bins:",
              choices = c(10, 20, 35, 50), selected = 20),
  
  sliderInput("bw_adjust", label = "Bandwidth adjustment:",
              min = 0.2, max = 2, value = 1, step = 0.2)
)

renderPlot({
  hist(faithful$eruptions, probability = TRUE, breaks = as.numeric(input$n_breaks),
       xlab = "Duration (minutes)", main = "Geyser eruption duration")
  
  dens <- density(faithful$eruptions, adjust = input$bw_adjust)
  lines(dens, col = "blue")
})
```

## Embedded Application

It's also possible to embed an entire Shiny application within an R Markdown document using the `shinyAppDir` function. This example embeds a Shiny application located in another directory:

```{r tabsets, echo=FALSE}
shinyAppDir(
  system.file("examples/06_tabsets", package = "shiny"),
  options = list(
    width = "100%", height = 550
  )
)
```

Note the use of the `height` parameter to determine how much vertical space the embedded application should occupy.

You can also use the `shinyApp` function to define an application inline rather then in an external directory.

In all of R code chunks above the `echo = FALSE` attribute is used. This is to prevent the R code within the chunk from rendering in the document alongside the Shiny components.



