---
title: "Seed mass index in FWS refuges"
author: "Emilio A. Laca"
date: "13 Feb 2022"
output: html_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)
source("SimMassPerAreaFun.R")
source("SimPropAreaFun.R")
source("SimMassBySuLITFun.R")
library(tidyverse)
library(knitr)
library(kableExtra)

```

## Perform simulations and calculate results for all years

```{r files, include=FALSE, echo=FALSE}

vpcp_path_2019 <- "inputFiles/vpcp2019.rds"
vpcp_path_2020 <- "inputFiles/vpcp2020.rds"
vpcp_path_2021 <- "inputFiles/vpcp2021.rds"

qdt_path_2019 <- "inputFiles/qdt_2019.rds"
qdt_path_2020 <- "inputFiles/qdt_2020.rds"
qdt_path_2021 <- "inputFiles/qdt_2021.rds"

```

```{r sims, include=FALSE, echo=FALSE}

set.seed(141416)

shmi_2019 <- sim_mass_su_lit(.vpcp_path = vpcp_path_2019,
                             .qdt_path = qdt_path_2019,
                             .nsim = 4000)

write_rds(shmi_2019, "outputFiles/shmi_2019.rds")


set.seed(141417)

shmi_2020 <- sim_mass_su_lit(.vpcp_path = vpcp_path_2020,
                             .qdt_path = qdt_path_2020,
                             .nsim = 4000)

write_rds(shmi_2020, "outputFiles/shmi_2020.rds")


set.seed(141418)

shmi_2021 <- sim_mass_su_lit(.vpcp_path = vpcp_path_2021,
                             .qdt_path = qdt_path_2021,
                             .nsim = 4000)

write_rds(shmi_2021, "outputFiles/shmi_2021.rds")

```

## Cumulative seed head mass index

Caption:

Cumulative seed head mass index for each refuge in year yyyy plotted against the corresponding cumulative area of subunits. Downward error bars show the lower confidence limit for a "right-side CI": 90% of the sampling distribution is above the lower extreme of the error bar. Solid error bars indicate that the distance from the estimate to the lower CI extreme is less than 25% of the estimate. Dashed bars indicate that precision is less than that necessary to achieve sampling goals.

```{r cum_shmi_plots, echo=FALSE}

cum_tot <- bind_rows(
  shmi_2019$cum_mass_LIT %>% mutate(year = 2019),
  shmi_2020$cum_mass_LIT %>% mutate(year = 2020),
  shmi_2021$cum_mass_LIT %>% mutate(year = 2021)
)

ggplot(data = cum_tot,
       aes(
         y = cum_tot_ton,
         x = cum_area_ha,
         groups = LIT,
         color = LIT
       )
) +
  geom_line(size = 1.0) +
  geom_errorbar(
    aes(
      ymin = ctt_RightCI90_lwr,
      ymax = cum_tot_ton,
      linetype = smpl_obj
    )
  ) +
  ylab("Cumulative seed head mass index (1000 kg or ton)") +
  xlab("Cumulative area (ha)") +
  geom_text(x = 350, y = 40, label = "Equal scales", color = "black") +
  facet_grid(rows = vars(year))


ggplot(data = cum_tot,
       aes(
         y = cum_tot_ton,
         x = cum_area_ha,
         groups = LIT,
         color = LIT
       )
) +
  geom_line(size = 1.0) +
  geom_errorbar(
    aes(
      ymin = ctt_RightCI90_lwr,
      ymax = cum_tot_ton,
      linetype = smpl_obj
    )
  ) +
  ylab("Cumulative seed head mass index (1000 kg or ton)") +
  xlab("Cumulative area (ha)") +
  geom_text(x = 350, y = 40,
            label = "Different scales",
            color = "black") +
  facet_grid(rows = vars(year),
             scales = "free")

```

Prepare table headers and function

```{r table_hds, echo=FALSE}

tbl_hds1 <- c("Unit",
             "Mean",
             "SE",
             "CI90lwr",
             "CI90upr",
             "RtCI90_lwr",
             "MOE/Mean",
             "Objective")

tbl_hds2 <- tbl_hds1[1:6]

# mk_kbl <- function(level = 1, yr = 2019, spp = "all") {
#   if (
#     !(level %in% c(1, 2)) | !(yr %in% c(2019, 2020, 2021)) 
#   ) {
#     stop("level must be 1 for subunits or 2 for refuge, and yr has to be 2019, 2020 or 2021.")
#   }
#   
#   # TBD ...
#   return( "Hello!")
# }

```

## Seed head mass index per area for units

All tables have a column for sampling objective based on data from the table. These are just for general information and are not being proposed as tests for the sampling objective. The only test to be considered is in the figures: if the error bar is dashed, the corresponding subunit does not meet precision objectives.

### All species together

Caption:

```{r shmi_tables_su_all_sp, echo=FALSE}

shmi_2019$tot_mass_su %>%
  set_names(nm = tbl_hds1) %>%
  kable(digits = c(0, 0, 1, 0, 0, 0, 2, 0),
    caption = "Average seed head mass index for all species in 2019 (g/m2)") %>%
  kable_styling(full_width = FALSE)

shmi_2020$tot_mass_su %>%
  set_names(nm = tbl_hds1) %>%
  kable(digits = c(0, 0, 1, 0, 0, 0, 2, 0),
    caption = "Average seed head mass index for all species in 2020 (g/m2)") %>%
  kable_styling(full_width = FALSE)


shmi_2021$tot_mass_su %>%
  set_names(nm = tbl_hds1) %>%
  kable(digits = c(0, 0, 1, 0, 0, 0, 2, 0),
    caption = "Average seed head mass index for all species in 2021 (g/m2)") %>%
  kable_styling(full_width = FALSE)

```

### Smartweed

Caption:

```{r shmi_tables_su_sw, echo=FALSE}

shmi_2019$sw_mass_m2 %>%
  set_names(nm = tbl_hds1) %>%
  kable(digits = c(0, 0, 1, 0, 0, 0, 2, 0),
    caption = "Average seed head mass index for Smartweed in 2019 (g/m2)") %>%
  kable_styling(full_width = FALSE)

shmi_2020$sw_mass_m2 %>%
  set_names(nm = tbl_hds1) %>%
  kable(digits = c(0, 0, 1, 0, 0, 0, 2, 0),
    caption = "Average seed head mass index for Smartweed in 2020 (g/m2)") %>%
  kable_styling(full_width = FALSE)


shmi_2021$sw_mass_m2 %>%
  set_names(nm = tbl_hds1) %>%
  kable(digits = c(0, 0, 1, 0, 0, 0, 2, 0),
    caption = "Average seed head mass index for Smartweed in 2021 (g/m2)") %>%
  kable_styling(full_width = FALSE)

```


### Swamp Timothy

Caption:

```{r shmi_tables_su_st, echo=FALSE}

shmi_2019$st_mass_m2 %>%
  set_names(nm = tbl_hds1) %>%
  kable(digits = c(0, 0, 1, 0, 0, 0, 2, 0),
    caption = "Average seed head mass index for Swamp Timothy in 2019 (g/m2)") %>%
  kable_styling(full_width = FALSE)

shmi_2020$st_mass_m2 %>%
  set_names(nm = tbl_hds1) %>%
  kable(digits = c(0, 0, 1, 0, 0, 0, 2, 0),
    caption = "Average seed head mass index for Swamp Timothy in 2020 (g/m2)") %>%
  kable_styling(full_width = FALSE)


shmi_2021$st_mass_m2 %>%
  set_names(nm = tbl_hds1) %>%
  kable(digits = c(0, 0, 1, 0, 0, 0, 2, 0),
    caption = "Average seed head mass index for Swamp Timothy in 2021 (g/m2)") %>%
  kable_styling(full_width = FALSE)

```


### Watergrass

Caption:

```{r shmi_tables_su_wg, echo=FALSE}

shmi_2019$wg_mass_m2 %>%
  set_names(nm = tbl_hds1) %>%
  kable(digits = c(0, 0, 1, 0, 0, 0, 2, 0),
    caption = "Average seed head mass index for Watergrass in 2019 (g/m2)") %>%
  kable_styling(full_width = FALSE)

shmi_2020$wg_mass_m2 %>%
  set_names(nm = tbl_hds1) %>%
  kable(digits = c(0, 0, 1, 0, 0, 0, 2, 0),
    caption = "Average seed head mass index for Watergrass in 2020 (g/m2)") %>%
  kable_styling(full_width = FALSE)


shmi_2021$wg_mass_m2 %>%
  set_names(nm = tbl_hds1) %>%
  kable(digits = c(0, 0, 1, 0, 0, 0, 2, 0),
    caption = "Average seed head mass index for Watergrass in 2021 (g/m2)") %>%
  kable_styling(full_width = FALSE)

```


## Seed head mass index per area for refuges

### All species

```{r shmi_table_lit_all_spp, echo=FALSE}

shmi_2019$lit_stats %>%
  dplyr::select(
    LIT,
    starts_with("tot")
  ) %>%
  set_names(nm = tbl_hds2) %>%
  kable(digits = c(0, 0, 1, 0, 0, 0),
    caption = "Average seed head mass index for all species in 2019 (g/m2)") %>%
  kable_styling(full_width = FALSE)

shmi_2020$lit_stats %>%
  dplyr::select(
    LIT,
    starts_with("tot")
  ) %>%
  set_names(nm = tbl_hds2) %>%
  kable(digits = c(0, 0, 1, 0, 0, 0),
    caption = "Average seed head mass index for all species in 2020 (g/m2)") %>%
  kable_styling(full_width = FALSE)

shmi_2021$lit_stats %>%
  dplyr::select(
    LIT,
    starts_with("tot")
  ) %>%
  set_names(nm = tbl_hds2) %>%
  kable(digits = c(0, 0, 1, 0, 0, 0),
    caption = "Average seed head mass index for all species in 2021 (g/m2)") %>%
  kable_styling(full_width = FALSE)

```

### Smartweed

```{r shmi_table_lit_sw, echo=FALSE}

shmi_2019$lit_stats %>%
  dplyr::select(
    LIT,
    starts_with("sw")
  ) %>%
  set_names(nm = tbl_hds2) %>%
  kable(digits = c(0, 0, 1, 0, 0, 0),
    caption = "Average seed head mass index for Smartweed in 2019 (g/m2)") %>%
  kable_styling(full_width = FALSE)

shmi_2020$lit_stats %>%
  dplyr::select(
    LIT,
    starts_with("sw")
  ) %>%
  set_names(nm = tbl_hds2) %>%
  kable(digits = c(0, 0, 1, 0, 0, 0),
    caption = "Average seed head mass index for Smartweed in 2020 (g/m2)") %>%
  kable_styling(full_width = FALSE)

shmi_2021$lit_stats %>%
  dplyr::select(
    LIT,
    starts_with("sw")
  ) %>%
  set_names(nm = tbl_hds2) %>%
  kable(digits = c(0, 0, 1, 0, 0, 0),
    caption = "Average seed head mass index for Smartweed in 2021 (g/m2)") %>%
  kable_styling(full_width = FALSE)

```


### Swamp Timothy

```{r shmi_table_lit_st, echo=FALSE}

shmi_2019$lit_stats %>%
  dplyr::select(
    LIT,
    starts_with("st")
  ) %>%
  set_names(nm = tbl_hds2) %>%
  kable(digits = c(0, 0, 1, 0, 0, 0),
    caption = "Average seed head mass index for Swamp Timothy in 2019 (g/m2)") %>%
  kable_styling(full_width = FALSE)

shmi_2020$lit_stats %>%
  dplyr::select(
    LIT,
    starts_with("st")
  ) %>%
  set_names(nm = tbl_hds2) %>%
  kable(digits = c(0, 0, 1, 0, 0, 0),
    caption = "Average seed head mass index for Swamp Timothy in 2020 (g/m2)") %>%
  kable_styling(full_width = FALSE)

shmi_2021$lit_stats %>%
  dplyr::select(
    LIT,
    starts_with("st")
  ) %>%
  set_names(nm = tbl_hds2) %>%
  kable(digits = c(0, 0, 1, 0, 0, 0),
    caption = "Average seed head mass index for Swamp Timothy in 2021 (g/m2)") %>%
  kable_styling(full_width = FALSE)

```

### Watergrass

```{r shmi_table_lit_wg, echo=FALSE}

shmi_2019$lit_stats %>%
  dplyr::select(
    LIT,
    starts_with("wg")
  ) %>%
  set_names(nm = tbl_hds2) %>%
  kable(digits = c(0, 0, 1, 0, 0, 0),
    caption = "Average seed head mass index for Watergrass in 2019 (g/m2)") %>%
  kable_styling(full_width = FALSE)

shmi_2020$lit_stats %>%
  dplyr::select(
    LIT,
    starts_with("wg")
  ) %>%
  set_names(nm = tbl_hds2) %>%
  kable(digits = c(0, 0, 1, 0, 0, 0),
    caption = "Average seed head mass index for Watergrass in 2020 (g/m2)") %>%
  kable_styling(full_width = FALSE)

shmi_2021$lit_stats %>%
  dplyr::select(
    LIT,
    starts_with("wg")
  ) %>%
  set_names(nm = tbl_hds2) %>%
  kable(digits = c(0, 0, 1, 0, 0, 0),
    caption = "Average seed head mass index for Watergrass in 2021 (g/m2)") %>%
  kable_styling(full_width = FALSE)

```

