---
title: "Calculation of Timothy and Watergrass seed abundance.Summer 2020"
author: "Emilio A. Laca"
date: "`r Sys.Date()`"
output: 
  html_document: 
    toc: yes
    number_sections: yes
editor_options: 
  chunk_output_type: console
---
<!--  Add code to remove the subunits that have all zeros. -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(car)
library(boot)
library(emmeans)
library(rsample)
library(MASS)
library(ggforce)
library(knitr)
library(kableExtra)

```

## Names of variables 

Whenever a different name is encountered for these variables, it will be corrected accordingly.

1. Refuge (for spelled-out refuge names, do not use as a key field)
2. LIT (use as the key field)
3. Unit_Name
4. Subunit_Name
5. subunit_ID (paste(LIT, Unit_Name, Subunit_Name, sep = "_"))
6. group (paste(LIT, Unit_Name, Subunit_Name, Stratum, sep = "_"))
7. Stratum ("a.Low", "b.Med", "c.High")
8. LIT_Strat  (factor(paste(LIT, Stratum, sep = "_")))


## Read RDS's

```{r rds_Files}

# Data to create model for ST: mass = f(seed no., seed length)
# From st_model_creation.R
st_all_noO <- readRDS("st_all_noO.rds") %>%
  rename_with(~gsub(".", "_", .x, fixed = TRUE))

# Data to create model for ST: mass = f(seed no., seed length)
# From wg_model_creation.R
wgdat <- readRDS("wgdat.rds") %>%
  rename_with(~gsub(".", "_", .x, fixed = TRUE))

# # Data from vantage points and circle plots from AreaPerSubunitPerStratum2019.R
# vpcp2 <- readRDS("/Users/emilioalaca/Google Drive/Documents/PROJECTS/RefugesIMKaylene/WaterbirdHabitat2017/WHAPeal/WHAPDataAnalysis/vpcp2.rds")

```

## Models to go from seed dimensions to mass

```{r ST_len2mass}

ST_len2mass <- lm(log(mass_mg) ~ log(avg_length_mm) + log(no_sh),
                  data = st_all_noO)

```

```{r WG_len2mass}

WG_len2mass <- lm(log(mass_per_sh_mg  + 20) ~ sh_len_cm +
                    emerged +
                    F2T_cm,
                  data = wgdat,
                  na.action = na.exclude)

```

## Prepare utility info

```{r utils}

# List of columns with seed head length and F2T--
length_col <- c("seed_head1_Length",
                "seed_head2_Length",
                "seed_head3_Length",
                "seed_head4_Length",
                "seed_head5_Length")

F2T_col <- c("seed_head1_F2TLength",
             "seed_head2_F2TLength",
             "seed_head3_F2TLength",
             "seed_head4_F2TLength",
             "seed_head5_F2TLength")

```

## Read data files

```{r prep_Files}

# Vantage polygon information =============================
vp2020c <- read.csv("vp2020c.csv")

vps <- vp2020c %>%
  mutate(level = "VP",
         temp_p = Watergrass_Low +
           Watergrass_Med +
           Watergrass_High +
           Timothy_Low +
           Timothy_Med +
           Timothy_High +
           NoFoodCover,
         p_wg_a.Low = Watergrass_Low / temp_p,
         p_wg_b.Med = Watergrass_Med / temp_p,
         p_wg_c.High = Watergrass_High / temp_p,
         p_st_a.Low = Timothy_Low / temp_p,
         p_st_b.Med = Timothy_Med / temp_p,
         p_st_c.High = Timothy_High / temp_p,
         p_noFood = NoFoodCover / temp_p,
         subunit_ID = factor(paste(LIT, Unit_Name, Subunit_Name, sep = "_"))) %>%
  dplyr::select(LIT,
                Unit_Name,
                Subunit_Name,
                subunit_ID,
                level,
                AreaVisible,
                starts_with("p_"))

# Circle plot information =============================
cp2020c <- read.csv("cp2020c.csv")


cps <- cp2020c %>%
  mutate(level = "CP",
         temp_p = Watergrass_Low +
           Watergrass_Med +
           Watergrass_High +
           Timothy_Low +
           Timothy_Med +
           Timothy_High,
         p_noFood = max(1 - temp_p, 0),
         p_wg_a.Low = Watergrass_Low / (p_noFood + temp_p),
         p_wg_b.Med = Watergrass_Med/ (p_noFood + temp_p),
         p_wg_c.High = Watergrass_High/ (p_noFood + temp_p),
         p_st_a.Low = Timothy_Low/ (p_noFood + temp_p),
         p_st_b.Med  = Timothy_Med/ (p_noFood + temp_p),
         p_st_c.High = Timothy_High/ (p_noFood + temp_p),
         p_noFood = p_noFood / (p_noFood + temp_p),
         subunit_ID = factor(paste(LIT, Unit_Name, Subunit_Name, sep = "_")),
         AreaVisible = 0.1746678) %>%
  dplyr::select(LIT,
                Unit_Name,
                Subunit_Name,
                subunit_ID,
                level,
                AreaVisible,
                starts_with("p_"))

# Management unit information =============================
mu2020c <- read.csv("mu2020c.csv")


mus <- mu2020c %>% # management units
  dplyr::select(LIT,
                Unit_Name,
                Subunit_Name,
                su.area_ac = Acreage) %>%
  mutate(subunit_ID = paste(LIT, Unit_Name, Subunit_Name, sep = "_"))

vpcp <- bind_rows(vps, cps) %>%
  full_join(mus) %>%
  na.omit()

# Quadrat information ====================================
qdt_st <- read.csv("qdt2020c.csv") %>%
  filter(CommonName == "Timothy" & Stratum != "") %>%
  mutate(avg_length_mm = 10 * rowMeans(.[, length_col], na.rm = TRUE)) %>%
  mutate(qdt_m2 = ifelse(Quadrat_Size == "15x15", 0.15^2,
                         ifelse(Quadrat_Size == "5x5", 0.05^2, 0.10)),
         n_seed_head_m2 = n_seed_heads / qdt_m2,
         no_sh = n_seed_head_m2) %>%
  dplyr::select(LIT,
                Unit_Name,
                Subunit_Name,
                Stratum,
                avg_length_mm,
                n_seed_head_m2,
                no_sh) %>%
  na.omit() %>%
  mutate(mass_g_m2 = unname(exp(predict(ST_len2mass, newdata = .)) / 1000),
         mass_g_m2 = ifelse(n_seed_head_m2 == 0, 0, mass_g_m2)) %>% # making sure it handles quadrats where no_sh == 0
  dplyr::filter(mass_g_m2 < 2000 & mass_g_m2 != 0) %>%
  mutate(Stratum = dplyr::recode(Stratum,
                                 `Low` = "a.Low",
                                 `Med` = "b.Med",
                                 `High` = "c.High"),
         Stratum = factor(as.character(Stratum)),
         subunit_ID = paste(LIT, Unit_Name, Subunit_Name, sep = "_"),
         LIT_Strat = factor(paste(LIT, Stratum, sep = "_"))) %>%
  arrange(LIT, Unit_Name, Subunit_Name, Stratum)

attributes(qdt_st)$na.action <- NULL

str(qdt_st)

# Proportion of area estimated from field

```

```{r wg_qdt}

# Quadrat data from field ========
qdt_wg <- read.csv("qdt2020c.csv") %>%
  filter(CommonName == "Watergrass" & Stratum != "")

sh_lengths_cm_wg <- as.matrix(qdt_wg[, length_col]) %>% na_if(0)
sh_F2T_cm_wg <- as.matrix(qdt_wg[, F2T_col]) %>% na_if(0)
sh_emerged_wg <- sh_F2T_cm_wg / sh_lengths_cm_wg

wgc <- unname(coef(WG_len2mass))

sh_mass_mg_mat <- exp(wgc[1] +
                        wgc[2] * sh_lengths_cm_wg +
                        wgc[3] * sh_emerged_wg +
                        wgc[4] * sh_F2T_cm_wg) - 20

qdt_wg <- qdt_wg %>%
  mutate(qdt_m2 = ifelse(Quadrat_Size == "15x15", 0.15^2,
                         ifelse(Quadrat_Size == "5x5", 0.05^2, 0.10)),
         n_seed_head_m2 = n_seed_heads / qdt_m2,
         sh_mass_mg = rowMeans(sh_mass_mg_mat, na.rm = TRUE),
         sh_mass_mg = ifelse(is.nan(sh_mass_mg), NA, sh_mass_mg),
         avg_length_mm = 10 * rowMeans(sh_lengths_cm_wg, na.rm = TRUE),
         mass_g_m2 = n_seed_head_m2 * sh_mass_mg / 1000) %>%
  filter(mass_g_m2 < 400) %>%
  dplyr:::select(LIT,
                 Unit_Name,
                 Subunit_Name,
                 Stratum,
                 avg_length_mm,
                 n_seed_head_m2,
                 sh_mass_mg,
                 mass_g_m2) %>%
  na.omit() %>%
  mutate(Stratum = dplyr::recode(Stratum,
                                 `Low` = "a.Low",
                                 `Med` = "b.Med",
                                 `High` = "c.High"),
         Stratum = factor(as.character(Stratum)),
         subunit_ID = paste(LIT, Unit_Name, Subunit_Name, sep = "_"),
         LIT_Strat = factor(paste(LIT, Stratum, sep = "_"))) %>%
  arrange(LIT, Unit_Name, Subunit_Name, Stratum)

attributes(qdt_wg)$na.action <- NULL

str(qdt_wg)

```

## Area per subunit

```{r su_areas}

su_areas <- mu2020c %>%
  dplyr::select(LIT,
                Unit_Name,
                Subunit_Name,
                area_ac = Acreage) %>%
  mutate(area_ha = area_ac / 2.47)

```

## Estimation of proportions of areas

Proportion of areas

See http://www.analyticalgroup.com/download/WEIGHTED_MEAN.pdf for var(weighted average) and AREA FRAME DESIGN.pdf

```{r p_areaAverages}

p_Area <- vpcp %>%
  group_by(subunit_ID) %>%
  summarise(across(starts_with("p_"), ~weighted.mean(.x, AreaVisible))) %>%
  arrange(subunit_ID)

# Uncorrected variance of the weighted average ========
v_Area <- vpcp %>% group_by(subunit_ID) %>% # see WEIGHTED_MEAN.pdf
        summarise(across(starts_with("p_"),
                         ~ var(.x) * sum(AreaVisible^2) / (sum(AreaVisible)^2))) %>%
  arrange(subunit_ID) # %>%
#  na.omit()

# Finite population correction factor ============
fpcf <- vpcp %>%
  group_by(subunit_ID) %>%
  summarise(AreaSampled_ac = sum(AreaVisible)) %>%
  mutate(subunit_ID2 = subunit_ID) %>%
  separate(subunit_ID2, sep = "_", into = c("LIT", "Unit_Name", "Subunit_Name")) %>%
  left_join(su_areas) %>%
  mutate(fpcf = 1 - AreaSampled_ac / area_ac,
         fpcf = ifelse(fpcf <= 0, 0.10, fpcf)) %>%
  dplyr::select(subunit_ID, fpcf) %>%
  arrange(subunit_ID)


# Standard error of weighted averages corrected for finite population =======
se_Area_c <- v_Area %>%
  full_join(fpcf) %>%
  mutate(across(starts_with("p_"), ~ sqrt(.x * fpcf))) %>%
  arrange(subunit_ID) # %>%
#  na.omit()

v_Area_c <- se_Area_c %>%
  mutate(across(starts_with("p_"), ~ .x^2)) %>%
  arrange(subunit_ID)

```

## Function to estimate mass per unit area in each stratum

```{r st.modelStrata}

# Use robust linear model
st_rlm3 <- rlm(log(mass_g_m2 + 5) ~ LIT_Strat,
              na.action = na.exclude,
              maxit = 40,
             data = qdt_st)

```

```{r wg.modelStrata}

with(qdt_wg, table(LIT, Stratum)) # only one quadrat in KRN; remove

qdt_wg <- qdt_wg %>% # Assumes there is no Watergrass in KRN
  filter(LIT != "KRN") %>%
  mutate(LIT = factor(LIT),
         LIT_Strat = factor(LIT_Strat))

# Use robust linear model. LIT and the interaction were NS
wg_rlm3 <- rlm(log(mass_g_m2 + 0.5) ~ LIT_Strat,
              na.action = na.exclude,
              maxit = 40,
             data = qdt_wg)

```

Make a data frame with the seed mass per unit area for each stratum, and species (columns) for each subunit (rows).
`seed_mass_fun` is a function that takes the file `p_Area`, and the models for g.m2 by Subunit_Name and Stratum ("st_rlm3" and "wg_rlm3") as arguments and yields the seed mass of st and wg per unit area of each subunit, and the area of each subunit. The resulting file can be used to calculate refuge-level yields.

```{r seed_mass_fun}

seed_mass_fun <- function(.p_Area, .qdt_st, .qdt_wg) {
  tryCatch({
    
    # make a model for mass per area at quadrat level for each species,
    # refuge and stratum
    .st_rlm3 <- rlm(log(mass_g_m2 + 0.5) ~ LIT_Strat, # make model for st
                    na.action = na.exclude,
                    maxit = 40,
                    data = .qdt_st)
    .wg_rlm3 <- rlm(log(mass_g_m2 + 5) ~ LIT_Strat, # make model for wg
                    na.action = na.exclude,
                    maxit = 40,
                    data = .qdt_wg)
    
    # create file to predict mass/area by stratum & subunit
    result <- .p_Area %>% 
      pull(subunit_ID) %>% # set of subunits for which there is % area estimates
      # create a df with all combinations of stratum and subunits present in p_Area
      expand_grid(subunit_ID = .,
                  Stratum = factor(c("a.Low", "b.Med", "c.High"))) %>%
      mutate(su = subunit_ID) %>% # add su column to recreate ID columns
      separate(su, sep = "_", into = c("LIT", "Unit_Name", "Subunit_Name")) %>%
      # create a LIT_Strat column to be able predict from .st_rlm3 and.wg_rlm3
      # first a oLIT_Strat with all rows is created
      mutate(oLIT_Strat = factor(paste(LIT, Stratum, sep = "_"))) %>%
      # make LIT_Strat col with nonNA values only for cases present in .st_rlm3
      mutate(LIT_Strat = factor(ifelse(oLIT_Strat %in% .st_rlm3$xlevels[[1]],
                                       as.character(oLIT_Strat),
                                       NA))) %>%
      # use current LIT_Strat to predict mass per area @ quadrat scale for st
      mutate(st_g_m2 = unname(exp(predict(.st_rlm3, newdata = .)) - 5),
      # make LIT_Strat col with nonNA values only for cases present in .wg_rlm3
             LIT_Strat = factor(ifelse(oLIT_Strat %in% .wg_rlm3$xlevels[[1]],
                                       as.character(oLIT_Strat),
                                       NA))) %>%
      # use current LIT_Strat to predict mass per area @ quadrat scale for wg
      mutate(wg_g_m2 =unname(exp(predict(.wg_rlm3, newdata = .)) - 0.5),
      # restore full LIT_Strat column
             LIT_Strat = oLIT_Strat,
      # Subunits present p_Area that have area estimates for a species and
      # stratum not represented in the species _rlm3 models are given 0 mass
      # This may result in an error and has to be avoided by making sure that
      # there for each refuge and species there are quadrats for all combinations
      # observed either at the VP or CP levels.
             st_g_m2 = replace_na(st_g_m2, 0),
             wg_g_m2 = replace_na(wg_g_m2, 0)) %>%
      dplyr::select(LIT,
                    Unit_Name,
                    Subunit_Name,
                    subunit_ID,
                    Stratum,
                    st_g_m2,
                    wg_g_m2) %>%
      # Strata for each species are separated into columns to calculate mass
      # per area in all 3 strata per species
      pivot_wider(names_from = Stratum,
                  values_from = c(st_g_m2, wg_g_m2),
                  values_fill = 0) %>%
      # join with proportion of area to weight strata by area
      full_join(.p_Area, by = c("subunit_ID")) %>%
      # weighted averages over the whole subunit per species
      # areas without the species (0 mass) are implicit as the sum of
      # all p_st and p_wg is less than 1 when there are no-food areas.
      mutate(wg_g_m2 = rowSums(dplyr::select(., starts_with("wg")) *
                                 dplyr::select(., starts_with("p_wg"))),
             st_g_m2 = rowSums(dplyr::select(., starts_with("st")) *
                                 dplyr::select(., starts_with("p_st"))),
             tot_g_m2 = wg_g_m2 + st_g_m2) %>%
      full_join(su_areas, by = c(c("LIT", "Unit_Name", "Subunit_Name"))) %>%
      dplyr::select(-starts_with("p_"), -area_ac) %>%
      na.omit()
    return(result)},
    error = function(e) return(NULL)
  )
}
```

## Mass per subunit and refuge

### Mass per subunit is given as average g/m2 over the whole subunit area. 

```{r mass_per_subunit}

qdt_mass_g_m2 <- seed_mass_fun(p_Area, qdt_st, qdt_wg) %>%
  arrange(subunit_ID) %>%
  na.omit()

attr(qdt_mass_g_m2, "notes") <- "These estimations exclude the uncertainty due to the proportion of area covered by each stratum in each unit, and the uncertainty of the model to convert seed number and seed length into mass."

str(qdt_mass_g_m2)
```


```{r table_, results='asis'}

qdt_mass_g_m2 %>%
  dplyr::select(LIT,
                Unit_Name,
                Subunit_Name,
                wg_g_m2,
                st_g_m2,
                tot_g_m2,
                area_ha) %>%
  kable(caption = "Average seed mass per subunit area. Total mass of seed estimated for the subunit divided by the total area of the subunit. These estimations exclude the uncertainty due to the proportion of area covered by each stratum in each unit, and the uncertainty of the model to convert seed number and seed length into mass.", 
        format = "html",
        digits = 1) %>%
  kable_styling(full_width = FALSE)

```


### Calculate mass per refuge (kg) based on original sample.

```{r mass_per_LIT}

mass_kg_LIT <- qdt_mass_g_m2 %>%
  group_by(LIT) %>%
  na.omit() %>%
  summarise(mass_kg_lit = sum(tot_g_m2 * area_ha * 10),
            tot_ha = sum(area_ha),
            .groups = "drop") %>%
  mutate(kg.ha = mass_kg_lit / tot_ha)

attr(mass_kg_LIT, "notes") <- "These estimations exclude the uncertainty due to the proportion of area covered by each stratum in each unit, and the uncertainty of the model to convert seed number and seed length into mass."

mass_kg_LIT %>%
  kable(digits = 0,
        format = "html",
        caption = "Total seed mass in each refuge, areas and seed mass per unit area. These estimations exclude the uncertainty due to the proportion of area covered by each stratum in each unit, and the uncertainty of the model to convert seed number and seed length into mass.") %>%
  kable_styling(full_width = FALSE)

```

## Mass estimates and distributions with quadrat variance.

Seed yield, confidence intervals. etc. for each subunit and refuge are calculated by resampling  yield per stratum for each subunit.

Using `tidymodels` and `rsample`.

```{r bootstrap_yield, warning=FALSE}

system.time(
  qdt_bt <- bootstraps(qdt_wg, # watergrass
                       times = 4000,
                       apparent = TRUE,
                       strata = LIT_Strat) %>%
    rename(qdt_wg = splits) %>%
    full_join(bootstraps(qdt_st,
                         times = 4000,
                         apparent = TRUE,
                         strata = LIT_Strat) %>%
                rename(qdt_st = splits), by = c("id")) %>% # timothy
    mutate(mass_su = map2(.x = qdt_st,
                          .y = qdt_wg,
                          ~seed_mass_fun(p_Area,
                                         .qdt_st = analysis(.x),
                                         .qdt_wg = analysis(.y))))
)

```

Create matrix of bootstrap estimates to calculate variances and covariances.

```{r booVcov}

  qdt_bt %>%
  dplyr::select(-starts_with("qdt_")) %>%
  unnest(cols = mass_su) %>%
  na.omit() %>%
  dplyr::select(id, subunit_ID, wg_g_m2, st_g_m2, tot_g_m2) %>%
  pivot_wider(names_from = subunit_ID, )


```



```{r mass_g_m2_Subunit_CI}

mass.su.CI <-
  qdt_bt %>%
  dplyr::select(-starts_with("qdt_")) %>%
  unnest(cols = mass_su) %>%
  na.omit() %>%
  dplyr::select(subunit_ID, wg_g_m2, st_g_m2, tot_g_m2, area_ha) %>%
  group_by(subunit_ID) %>%
  summarize(wg_q10 = quantile(wg_g_m2, 0.10, na.rm = TRUE),
            wg_mean = mean(wg_g_m2, na.rm = TRUE),
            wg_q90 = quantile(wg_g_m2, 0.90, na.rm = TRUE),
            st_q10 = quantile(st_g_m2, 0.10, na.rm = TRUE),
            st_mean = mean(st_g_m2, na.rm = TRUE),
            st_q90 = quantile(st_g_m2, 0.90, na.rm = TRUE),
            tot_q10 = quantile(tot_g_m2, 0.10, na.rm = TRUE),
            tot_mean = mean(tot_g_m2, na.rm = TRUE),
            tot_q90 = quantile(tot_g_m2, 0.90, na.rm = TRUE)) %>%
  arrange(subunit_ID) %>% na.omit()

attr(mass.su.CI, "notes") <- "These estimations exclude the uncertainty due to the proportion of area covered by each stratum in each unit, and the uncertainty of the model to convert seed number and seed length into mass."


print(mass.su.CI, n = Inf)

```


```{r mass_kg_LIT_CI}

mass_kg_LIT_CI <- qdt_bt %>%
  filter(unlist(map(mass_su, ~!is.null(.x)))) %>%
  mutate(mass_su = map(.x = mass_su, ~drop_na(.x)),
         mass_lit = map(.x = mass_su,
                        ~group_by(.x, LIT) %>%
                          summarise(mass_kg_lit = sum(tot_g_m2 * area_ha * 10),
                                    tot_ha = sum(area_ha),
                                    kg_ha = mass_kg_lit / tot_ha,
                                    .groups = "drop"))) %>%
  dplyr::select(mass_lit) %>%
  unnest(cols = mass_lit) %>%
  group_by(LIT) %>%
  summarise(mass_kg_lit_q10 = quantile(mass_kg_lit, 0.10, na.rm = TRUE),
            mass_kg_lit_q90 = quantile(mass_kg_lit, 0.90, na.rm = TRUE),
            mass_kg_lit_mean = mean(mass_kg_lit, na.rm = TRUE),
            kg_ha_q10 = quantile(kg_ha, 0.10, na.rm = TRUE),
            kg_ha_q90 = quantile(kg_ha, 0.90, na.rm = TRUE),
            kg_ha_mean = mean(kg_ha, na.rm = TRUE),
            tot_ha = mean(tot_ha, 0.10, na.rm = TRUE), .groups = "drop")

attr(mass_kg_LIT_CI, "notes") <- "These estimations exclude the uncertainty due to the proportion of area covered by each stratum in each unit, and the uncertainty of the model to convert seed number and seed length into mass."

```

## Mass estimates with quadrat and area variance
### Sum of products of independent variables

This approach is approximate because it assumes that composition variables are independent, but it includes the correction for finite population.

The formula used is:

$$ \sigma^2_{\sum(X_i Y_i)} = \sum{\left((\sigma^2_{X_i} + \mu^2_{X_i})(\sigma^2_{Y_i} + \mu^2_{Y_i}) - \mu^2_{X_i} \mu^2_{Y_i}\right)}$$

where $X_i$ are the values of mass per unit are for each stratum and the $Y_i$ are the proportions of each stratum in the subunit. The variances of $X_i$ are obtained from the bootstrapped data and the variances of $Y_i$ are given by the calculation of `v_Area_c`.

```{r area_qdt_var_included}

mass_var <-
  qdt_bt %>%
  dplyr::select(-starts_with("qdt_")) %>%
  filter(unlist(map(qdt_bt$mass_su, ~!is.null(.x)))) %>%
  mutate(
    mass_su =
      map(.x = mass_su,
          ~unique(.x %>%
                    dplyr::select(LIT, contains("_g_m2_"))))
  ) %>%
  unnest(cols = mass_su) %>%
  na.omit() %>%
  group_by(LIT) %>%
  summarise(across(contains("_g_m2_"), ~var(.x)),
            .groups = "drop") %>%
  arrange(LIT) %>% na.omit()

# Variance of total seed head mass per unit (wg and st together)

Xs2 <- v_Area_c %>%
  arrange(subunit_ID) %>%
  dplyr::select(p_st_a.Low, # variance of areas
                p_st_b.Med, 
                p_st_c.High,
                p_wg_a.Low,
                p_wg_b.Med, 
                p_wg_c.High)

Xm2 <- p_Area %>%
  arrange(subunit_ID) %>%
  dplyr::select(p_st_a.Low, # means of areas
                p_st_b.Med, 
                p_st_c.High,
                p_wg_a.Low,
                p_wg_b.Med, 
                p_wg_c.High) %>%
  `^`(2) # square everything

Ys2 <- qdt_mass_g_m2 %>%
  dplyr::select(-contains("_g_m2"), -area_ha) %>%
  left_join(mass_var) %>%
  arrange(subunit_ID) %>%
  dplyr::select(contains("_g_m2_"))

Ym2 <- (qdt_mass_g_m2 %>%
          arrange(subunit_ID) %>%
          dplyr::select(contains("_g_m2_")))^2

Xs2_st <- Xs2 %>% dplyr::select(contains("st_"))
Xm2_st <- Xm2 %>% dplyr::select(contains("st_"))
Ys2_st <- Ys2 %>% dplyr::select(contains("st_"))
Ym2_st <- Ym2 %>% dplyr::select(contains("st_"))

Xs2_wg <- Xs2 %>% dplyr::select(contains("wg_"))
Xm2_wg <- Xm2 %>% dplyr::select(contains("wg_"))
Ys2_wg <- Ys2 %>% dplyr::select(contains("wg_"))
Ym2_wg <- Ym2 %>% dplyr::select(contains("wg_"))


qdt_mass_g_m2 <- qdt_mass_g_m2 %>%
  mutate(tot_mass_m2_var =
           ((Xs2 + Xm2) *  (Ys2 + Ym2) - Xm2 * Ym2) %>%
           rowSums(),
         CI80lwr = tot_g_m2 + qnorm(0.10) * sqrt(tot_mass_m2_var),
         CI80upr = tot_g_m2 + qnorm(0.90) * sqrt(tot_mass_m2_var),
         st_mass_m2_var =
           ((Xs2_st + Xm2_st) *  (Ys2_st + Ym2_st) - Xm2_st * Ym2_st) %>%
           rowSums(),
         CI80lwr_st = st_g_m2 + qnorm(0.10) * sqrt(st_mass_m2_var),
         CI80upr_st = st_g_m2 + qnorm(0.90) * sqrt(st_mass_m2_var),
         wg_mass_m2_var =
           ((Xs2_wg + Xm2_wg) *  (Ys2_wg + Ym2_wg) - Xm2_wg * Ym2_wg) %>%
           rowSums(),
         CI80lwr_wg = wg_g_m2 + qnorm(0.10) * sqrt(wg_mass_m2_var),
         CI80upr_wg = wg_g_m2 + qnorm(0.90) * sqrt(wg_mass_m2_var))


qdt_mass_g_m2 %>%
  dplyr::select(LIT, Unit_Name, Subunit_Name, wg_g_m2, st_g_m2, CI80lwr, tot_g_m2, CI80upr) %>%
  mutate(delta_pcnt = round((100 * (CI80upr - CI80lwr) / 2) / tot_g_m2, 0)) %>%
  arrange(delta_pcnt) %>%
  print(n = Inf)

attributes(qdt_mass_g_m2)$notes <- "These estimates include the variance due to quadrat sampling and estimation of proportion of area covered by each stratum. They exclude the variance of the models to convert seed number and length into mass."

write.csv(qdt_mass_g_m2, "qdt_mass_g_m2_2020.txt")

```


## Graphical display of results =======

```{r CI_graphs}

# SHADE VERTICAL VIABILITY REGIONS
# Confidence intervals and estimates are from chunk `area_qdt_var_included`

class.bounds.g.m2 <- c(0, 50, 150, 250)

# Total mass per m2 in each subunit.

for (i in 1:length(unique(qdt_mass_g_m2$LIT))) {
  plot <- ggplot(data = qdt_mass_g_m2 %>% arrange(LIT, tot_g_m2),
                 aes(x = subunit_ID,
                     y = tot_g_m2,
                     ymin = CI80lwr,
                     ymax = CI80upr,
                     group = LIT)) +
#    xlim(0, 600) +
    geom_pointrange() +
    geom_hline(col = rep(c("red", "orange", "yellow", "green"),
                         length(unique(qdt_mass_g_m2$LIT))),
               yintercept = class.bounds.g.m2, linetype = 2) +
    xlab('Subunit_Name') + ylab("Total seed mass 80% CI (g/m^2)") +
    geom_errorbar(aes(ymin = CI80lwr, ymax = CI80upr), width = 0.5, cex = 1) +
    theme(plot.title = element_text(size = 16, face = "bold"),
          #        axis.text.y = element_blank(),
          axis.ticks.y = element_blank(),
          axis.text.x = element_text(face = "bold"),
          axis.title = element_text(size = 12, face = "bold"),
          strip.text.y = element_text(hjust = 0, vjust = 1, angle = 180, face = "bold")) +
    #  scale_y_log10() +
    coord_flip() +
    facet_wrap_paginate(~LIT,
                        strip.position = "left",
                        nrow = 1,
                        ncol = 1,
                        scales = "free_y",
                        page = i)
  print(plot)
}

# Timothy mass per m2 in each subunit.

for (i in 1:length(unique(qdt_mass_g_m2$LIT))) {
  plot <- ggplot(data = qdt_mass_g_m2 %>% arrange(LIT, tot_g_m2),
                 aes(x = subunit_ID,
                     y = st_g_m2,
                     ymin = CI80lwr_st,
                     ymax = CI80upr_st,
                     group = LIT)) +
    geom_pointrange() +
    geom_hline(col = rep(c("red", "orange", "yellow", "green"),
                         length(unique(qdt_mass_g_m2$LIT))),
               yintercept = class.bounds.g.m2, linetype = 2) +
    xlab('Subunit_Name') + ylab("Timothy seed mass 80% CI (g/m^2)") +
    geom_errorbar(aes(ymin = CI80lwr_st, ymax = CI80upr_st), width = 0.5, cex = 1) +
    theme(plot.title = element_text(size = 16, face = "bold"),
          #        axis.text.y = element_blank(),
          axis.ticks.y = element_blank(),
          axis.text.x = element_text(face = "bold"),
          axis.title = element_text(size = 12, face = "bold"),
          strip.text.y = element_text(hjust = 0, vjust = 1, angle = 180, face = "bold")) +
    #  scale_y_log10() +
    coord_flip() +
    facet_wrap_paginate(~LIT,
                        strip.position = "left",
                        nrow = 1,
                        ncol = 1,
                        scales = "free_y",
                        page = i)
  print(plot)
}

# Watergrass mass per m2 in each subunit.

for (i in 1:length(unique(qdt_mass_g_m2$LIT))) {
  plot <- ggplot(data = qdt_mass_g_m2 %>% arrange(LIT, tot_g_m2),
                 aes(x = subunit_ID,
                     y = wg_g_m2,
                     ymin = CI80lwr_wg,
                     ymax = CI80upr_wg,
                     group = LIT)) +
    geom_pointrange() +
    geom_hline(col = rep(c("red", "orange", "yellow", "green"),
                         length(unique(qdt_mass_g_m2$LIT))),
               yintercept = class.bounds.g.m2, linetype = 2) +
    xlab('Subunit_Name') + ylab("Watergrass seed mass 80% CI (g/m^2)") +
    geom_errorbar(aes(ymin = CI80lwr_wg, ymax = CI80upr_wg), width = 0.5, cex = 1) +
    theme(plot.title = element_text(size = 16, face = "bold"),
          #        axis.text.y = element_blank(),
          axis.ticks.y = element_blank(),
          axis.text.x = element_text(face = "bold"),
          axis.title = element_text(size = 12, face = "bold"),
          strip.text.y = element_text(hjust = 0, vjust = 1, angle = 180, face = "bold")) +
    #  scale_y_log10() +
    coord_flip() +
    facet_wrap_paginate(~LIT,
                        strip.position = "left",
                        nrow = 1,
                        ncol = 1,
                        scales = "free_y",
                        page = i)
  print(plot)
}

```



## Test of differences in mass/m.sq among units within refuge
```{r compareSubunits}
# Create function to do differences between all column pairs

compare.all.su <- function(mydf) {
  row.no <- 0
  outmydf <- data.frame(diff.name = character(),
                        mean.a = numeric(),
                        mean.b = numeric(),
                        mean.diff = numeric(),
                        var.a = numeric(),
                        var.b = numeric(),
                        sd.diff = numeric(),
                        tcalc = numeric(),
                        approx_p_value = numeric(),
                        stringsAsFactors = FALSE)
  
  for (i in 1:(nrow(mydf) - 1)) {
    for (j in (i + 1):nrow(mydf)) {
      row.no <- row.no + 1
      x1 <- paste(mydf$subunit_ID[j], mydf$subunit_ID[i], sep = " vs. ")
      x2 <- mydf$tot_g_m2[j]
      x3  <- mydf$tot_g_m2[i]
      x4 <- x2 - x3
      x5 <- mydf$tot_mass_m2_var[j]
      x6 <- mydf$tot_mass_m2_var[i]
      x7 <- sqrt(x5 + x6)
      x8 <- x4/x7
      x9 <- round(pnorm(q = abs(x8), lower.tail = FALSE), 9)
      outmydf[row.no, "diff.name"] <- x1
      outmydf[row.no, 2:9] <- c(x2, x3, x4, x5, x6, x7, x8, x9)
    }
  }
  return(outmydf)
}

compare.all.st <- function(mydf) {
  row.no <- 0
  outmydf <- data.frame(diff.name = character(),
                        mean.a = numeric(),
                        mean.b = numeric(),
                        mean.diff = numeric(),
                        var.a = numeric(),
                        var.b = numeric(),
                        sd.diff = numeric(),
                        tcalc = numeric(),
                        approx_p_value = numeric(),
                        stringsAsFactors = FALSE)
  
  for (i in 1:(nrow(mydf) - 1)) {
    for (j in (i + 1):nrow(mydf)) {
      row.no <- row.no + 1
      x1 <- paste(mydf$subunit_ID[j], mydf$subunit_ID[i], sep = " vs. ")
      x2 <- mydf$st_g_m2[j]
      x3  <- mydf$st_g_m2[i]
      x4 <- x2 - x3
      x5 <- mydf$st_mass_m2_var[j]
      x6 <- mydf$st_mass_m2_var[i]
      x7 <- sqrt(x5 + x6)
      x8 <- x4/x7
      x9 <- round(pnorm(q = abs(x8), lower.tail = FALSE), 9)
      outmydf[row.no, "diff.name"] <- x1
      outmydf[row.no, 2:9] <- c(x2, x3, x4, x5, x6, x7, x8, x9)
    }
  }
  return(outmydf)
}

compare.all.wg <- function(mydf) {
  row.no <- 0
  outmydf <- data.frame(diff.name = character(),
                        mean.a = numeric(),
                        mean.b = numeric(),
                        mean.diff = numeric(),
                        var.a = numeric(),
                        var.b = numeric(),
                        sd.diff = numeric(),
                        tcalc = numeric(),
                        approx_p_value = numeric(),
                        stringsAsFactors = FALSE)
  
  for (i in 1:(nrow(mydf) - 1)) {
    for (j in (i + 1):nrow(mydf)) {
      row.no <- row.no + 1
      x1 <- paste(mydf$subunit_ID[j], mydf$subunit_ID[i], sep = " vs. ")
      x2 <- mydf$wg_g_m2[j]
      x3  <- mydf$wg_g_m2[i]
      x4 <- x2 - x3
      x5 <- mydf$wg_mass_m2_var[j]
      x6 <- mydf$wg_mass_m2_var[i]
      x7 <- sqrt(x5 + x6)
      x8 <- x4/x7
      x9 <- round(pnorm(q = abs(x8), lower.tail = FALSE), 9)
      outmydf[row.no, "diff.name"] <- x1
      outmydf[row.no, 2:9] <- c(x2, x3, x4, x5, x6, x7, x8, x9)
    }
  }
  return(outmydf)
}

compare_su_mass <- qdt_mass_g_m2 %>%
  dplyr::select(LIT,
                subunit_ID,
                wg_g_m2,
                st_g_m2,
                tot_g_m2,
                wg_mass_m2_var,
                st_mass_m2_var,
                tot_mass_m2_var) %>%
  group_by(LIT) %>%
  nest() %>%
  mutate(tot_mass = map(data, ~compare.all.su(.x)),
         st_mass = map(data, ~ compare.all.st(.x)),
         wg_mass = map(data, ~ compare.all.wg(.x)))

## Comparisons for Total seed mass across subunits in each Refuge #########

for (lit in 1:length(unique(qdt_mass_g_m2$LIT))) {
compare_su_mass$tot_mass[[lit]] %>%
  dplyr::select(diff.name,
                mean.a,
                mean.b,
                approx_p_value) %>%
  as_tibble() %>%
  print(n = Inf)
}

# Output all tables as csv.

compare_su_mass %>%
  dplyr::select(st_mass) %>%
  unnest(cols = c(st_mass)) %>%
  write_csv("CompareSTMassWithinRefuge_2020.csv")

compare_su_mass %>%
  dplyr::select(wg_mass) %>%
  unnest(cols = c(wg_mass)) %>%
  write_csv("CompareWGMassWithinRefuge_2020.csv")

compare_su_mass %>%
  dplyr::select(tot_mass) %>%
  unnest(cols = c(tot_mass)) %>%
  write_csv("CompareTotalMassWithinRefuge_2020.csv")



```

```{r yieldComponents}

# Proportion of area in each subunit ====================

kable(p_Area,
      digits = c(0, rep(3, dim(p_Area)[2] - 1)),
      caption = "Proportion of area in each stratum by subunit") %>%
  kable_styling()

write.csv(p_Area, "../Output2020/p_Area2020.csv")

# Swamp Timothy ====================

st_qdt_means <- emmeans(st_rlm3, "LIT_Strat", type = "response") %>%
  as.tibble() %>%
  arrange(as.character(LIT_Strat)) %>%
  dplyr::select(-df) %>%
  rename(st_g_m2 = response)

kable(st_qdt_means,
      digits = c(0, 1, 2, 1, 1),
      caption = "Estimated g/m2 of seedheads for Swamp Timothy strata") %>%
  kable_styling(full_width = FALSE)

write.csv(st_qdt_means,
          "../Output2020/TimothyMassPerAreaByLITstratum2020.csv")

# Watergrass ====================

wg_qdt_means <- emmeans(wg_rlm3, "LIT_Strat", type = "response") %>%
  as.tibble() %>%
  arrange(as.character(LIT_Strat)) %>%
  dplyr::select(-df) %>%
  rename(wg_g_m2 = response)

kable(wg_qdt_means,
      digits = c(0, 1, 2, 1, 1),
      caption = "Estimated g/m2 of seedheads for Watergrass strata") %>%
  kable_styling(full_width = FALSE)

write.csv(wg_qdt_means,
          "../Output2020/WatergrassMassPerAreaByLITstratum2020.csv")

```

## Save objects for demo shiny app

```{r}
# Save estimates for plots and display

saveRDS(qdt_mass_g_m2, "../Output2020/qdt_mass_g_m2_2020.rds")

# Save comparisons between units within refuge

saveRDS(compare_su_mass, "../Output2020/compare_su_mass_2020.rds")

```



